---
title: "DVGfinder Analysis 2.0"
author: "RC"
date: "2025-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(car)
library(MASS)
library(purrr)
library(FSA)
```


Let's open the data and look at it.

```{r}
############################################################
## 0. LOAD BASE INPUT FILES
############################################################

consensus <- read.csv("DVGfinder_CONSENSUS_combined.csv", sep=',', header=TRUE)
#ml        <- read.csv("DVGfinder_FILTERED_ML_combined.csv", sep=',', header=TRUE)
#all       <- read.csv("DVGfinder_ALL_combined.csv", sep=',', header=TRUE)

coverage <- read_csv("coverage_pol.csv")   # will be used for filtering based on coverage
meta     <- read.csv("Metadata.csv", sep=',', header=TRUE)


output_sample <- "sample_summary.csv"
output_group  <- "groups_summary.csv"
```


```{r}
############################################################
## 1. PARAMETERS (adjust here)
############################################################

min_support <- 5              # Minimum read support for DVGs
coverage_threshold <- 10      # Median polymerase coverage threshold
polymerase_segments <- c("PB1","PB2","PA")

# Select which DVG table to analyze
dvg <- consensus    # or all / ml

############################################################
## 2. NORMALIZE SAMPLEID AND JOIN METADATA
############################################################

dvg  <- dvg  %>% mutate(SampleID = suppressWarnings(as.character(SampleID)))
meta <- meta %>% mutate(SampleID = suppressWarnings(as.character(SampleID)))

dvg <- left_join(dvg, meta, by = "SampleID")

dvg <- dvg %>%
  filter(Type == "HOM") %>% 
  filter(!SampleID %in% c("9770", "10318"))

############################################################
## 3. REORDER COLUMNS (METADATA FIRST)
############################################################

meta_cols   <- intersect(c("SampleID", setdiff(names(meta), "SampleID")), names(dvg))
segment_col <- "Segment"
dvg_cols    <- setdiff(names(dvg), c(meta_cols, segment_col))

dvg <- dvg[, c(meta_cols, segment_col, dvg_cols)]

if (!dir.exists("Outputs")) dir.create("Outputs")
stats_file <- "Outputs/dvg_stats_summary.txt"
file.create(stats_file)

############################################################
## 4. COVERAGE FILTERING (coverage_pol.csv)
############################################################

coverage <- coverage %>% mutate(SampleID = as.character(SampleID))
pol_cols <- grep("^(PB1|PB2|PA)\\.", colnames(coverage), value = TRUE)

sample_cov <- coverage %>%
  dplyr::mutate(
    median_pol_cov = apply(
      dplyr::select(., all_of(pol_cols)),
      1,
      median,
      na.rm = TRUE
    )
  ) %>%
  dplyr::select(SampleID, median_pol_cov)

good_samples <- sample_cov %>%
  filter(median_pol_cov >= coverage_threshold) %>%
  pull(SampleID)

dvg <- dvg %>%
  filter(SampleID %in% good_samples)

############################################################
## 5. FILTER DVG EVENTS (support, NA checks, length)
############################################################

dvg_filt <- dvg %>%
  filter(
    !is.na(BP),
    !is.na(RI),
    length_dvg > 0,
    mean_read_counts >= min_support,
    DVG_type != "Deletion_reverse"
  )


############################################################
## 5b. CLASSIFY IN-FRAME VS FRAMESHIFT (per event)
############################################################

dvg_filt <- dvg_filt %>%
  mutate(
    del_nt    = RI - BP + 1,
    in_frame  = if_else(del_nt %% 3 == 0, TRUE, FALSE, missing = FALSE),
    deletive  = grepl("Deletion", DVG_type, ignore.case = TRUE)
  )
# `deletive` lets you subset to true deletion events only if you want.

############################################################
## 6. HELPER FUNCTIONS
############################################################

shannon_entropy <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  p <- table(x) / length(x)
  -sum(p * log(p))
}

shannon_weighted_reads <- function(reads) {
  reads <- reads[is.finite(reads) & reads > 0]
  if (length(reads) == 0) return(NA_real_)
  p <- reads / sum(reads)
  -sum(p * log(p))
}

q90_safe <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_real_)
  as.numeric(quantile(x, 0.9, names = FALSE))
}

log_to_stats <- function(..., sep = " ") {
  cat(..., sep = sep, file = stats_file, append = TRUE)
  cat("\n", file = stats_file, append = TRUE)
}

log_block <- function(obj) {
  cat(capture.output(print(obj)), sep = "\n",
      file = stats_file, append = TRUE)
}

results_list <- list()

format_p <- function(p) {
  if (is.na(p)) return(NA_character_)
  if (p < 0.0001) return(format(p, scientific = TRUE, digits = 3))
  sprintf("%.4f", p)
}

############################################################
## 7. PER-SAMPLE METRICS (EVENT-LEVEL INPUT: dvg_filt)
############################################################

viral_load_est <- dvg_filt %>%
  filter(mean_read_counts > 0, mean_rpht > 0) %>%
  mutate(total_viral_reads_est = mean_read_counts * 1e5 / mean_rpht) %>%
  group_by(SampleID) %>%
  summarise(
    viral_reads_est = median(total_viral_reads_est, na.rm = TRUE),
    .groups = "drop"
  )

## 7a. per-sample, per-segment read-based polymerase fraction and frame stats

# Per-sample per-segment polymerase read fraction and frame counts
segment_frame_stats <- dvg_filt %>%
  group_by(across(all_of(meta_cols)), Segment) %>%
  summarise(
    segment_support          = sum(mean_read_counts, na.rm = TRUE),
    seg_inframe_reads        = sum(mean_read_counts[deletive & in_frame], na.rm = TRUE),
    seg_frameshift_reads     = sum(mean_read_counts[deletive & !in_frame], na.rm = TRUE),
    seg_inframe_count        = sum(deletive & in_frame, na.rm = TRUE),
    seg_frameshift_count     = sum(deletive & !in_frame, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(across(all_of(meta_cols))) %>%
  mutate(
    total_segment_support    = sum(segment_support, na.rm = TRUE),
    frac_segment_support     = if_else(total_segment_support > 0,
                                       segment_support / total_segment_support,
                                       NA_real_),
    frac_inframe_reads_seg   = if_else(segment_support > 0,
                                       seg_inframe_reads / segment_support,
                                       NA_real_),
    frac_inframe_events_seg  = if_else(
                                 (seg_inframe_count + seg_frameshift_count) > 0,
                                 seg_inframe_count / (seg_inframe_count + seg_frameshift_count),
                                 NA_real_
                               )
  ) %>%
  ungroup()

write.csv(
  segment_frame_stats,
  "Outputs/dvg_segment_frame_stats_per_sample.csv",
  row.names = FALSE
)

## 7b. per-sample collapsed metrics (what you had, plus new polymerase read fraction and frame totals)

sample_metrics <- dvg_filt %>%
  group_by(across(all_of(meta_cols))) %>%
  summarise(
    DVG_count        = n_distinct(cID_DI),

    mean_length      = mean(length_dvg),
    median_length    = median(length_dvg),
    p90_length       = q90_safe(length_dvg),
    sd_length        = sd(length_dvg),
    IQR_length       = IQR(length_dvg),

    total_support    = sum(mean_read_counts),
    sum_rpht         = sum(mean_rpht, na.rm = TRUE),

    shannon_weighted = shannon_weighted_reads(mean_read_counts),
    top1_fraction    = max(mean_read_counts) / sum(mean_read_counts),

    BP_entropy       = shannon_entropy(paste(Segment, BP, sep = ":")),

    # fraction of DVG *species* (unique DI events) in PB1/PB2/PA
    fraction_polymerase_species = mean(Segment %in% polymerase_segments),

    n_segments_with_dvg = n_distinct(Segment),

    # read-level polymerase support (THIS MUST BE NUMERIC)
    polymerase_support = sum(mean_read_counts[Segment %in% polymerase_segments], na.rm = TRUE),

    fraction_polymerase_reads = sum(mean_read_counts[Segment %in% polymerase_segments], na.rm = TRUE) /
                            sum(mean_read_counts, na.rm = TRUE),

    # In-frame vs frameshift totals (deletion events only)
    inframe_reads_total    = sum(mean_read_counts[deletive & in_frame], na.rm = TRUE),
    frameshift_reads_total = sum(mean_read_counts[deletive & !in_frame], na.rm = TRUE),

    inframe_events_total    = sum(deletive & in_frame),
    frameshift_events_total = sum(deletive & !in_frame),

    frac_inframe_reads_total = sum(mean_read_counts[deletive & in_frame], na.rm = TRUE) /
                               (sum(mean_read_counts[deletive & in_frame], na.rm = TRUE) +
                                sum(mean_read_counts[deletive & !in_frame], na.rm = TRUE)),

    frac_inframe_events_total = sum(deletive & in_frame) /
                                (sum(deletive & in_frame) + sum(deletive & !in_frame)),

    .groups = "drop"
  ) %>%
  left_join(viral_load_est, by = "SampleID") %>%
  mutate(
    normalized_richness_rpht =
      ifelse(is.finite(viral_reads_est) & viral_reads_est > 0,
             DVG_count / viral_reads_est * 1e5,
             NA_real_),

    normalized_abundance_rpht =
      ifelse(is.finite(viral_reads_est) & viral_reads_est > 0,
             total_support / viral_reads_est * 1e5,
             NA_real_)
  )


############################################################
## 8. ZERO-DVG SAMPLES
############################################################

sample_metrics_full <- all_samples %>%
  left_join(sample_metrics, by = meta_cols) %>%
  mutate(
    DVG_count = replace_na(DVG_count, 0),

    mean_length   = ifelse(DVG_count == 0, NA_real_, mean_length),
    median_length = ifelse(DVG_count == 0, NA_real_, median_length),
    p90_length    = ifelse(DVG_count == 0, NA_real_, p90_length),
    sd_length     = ifelse(DVG_count == 0, NA_real_, sd_length),
    IQR_length    = ifelse(DVG_count == 0, NA_real_, IQR_length),

    total_support = ifelse(DVG_count == 0, 0, total_support),
    sum_rpht      = ifelse(DVG_count == 0, 0, sum_rpht),

    shannon_weighted = ifelse(DVG_count == 0, NA_real_, shannon_weighted),
    top1_fraction    = ifelse(DVG_count == 0, NA_real_, top1_fraction),

    BP_entropy = ifelse(DVG_count == 0, NA_real_, BP_entropy),

    fraction_polymerase_species = ifelse(DVG_count == 0, NA_real_, fraction_polymerase_species),
    n_segments_with_dvg         = ifelse(DVG_count == 0, NA_real_, n_segments_with_dvg),

    polymerase_support    = ifelse(DVG_count == 0, 0, polymerase_support),
    fraction_polymerase_reads = ifelse(DVG_count == 0, NA_real_, fraction_polymerase_reads),

    inframe_reads_total     = ifelse(DVG_count == 0, 0, inframe_reads_total),
    frameshift_reads_total  = ifelse(DVG_count == 0, 0, frameshift_reads_total),
    inframe_events_total    = ifelse(DVG_count == 0, 0, inframe_events_total),
    frameshift_events_total = ifelse(DVG_count == 0, 0, frameshift_events_total),

    frac_inframe_reads_total  = ifelse(DVG_count == 0, NA_real_, frac_inframe_reads_total),
    frac_inframe_events_total = ifelse(DVG_count == 0, NA_real_, frac_inframe_events_total),

    viral_reads_est           = ifelse(DVG_count == 0, NA_real_, viral_reads_est),
    normalized_richness_rpht  = ifelse(DVG_count == 0, NA_real_, normalized_richness_rpht),
    normalized_abundance_rpht = ifelse(DVG_count == 0, NA_real_, normalized_abundance_rpht)
  )


write.csv(sample_metrics_full,
          "Outputs/dvg_sample_metrics.csv",
          row.names = FALSE)

############################################################
## 9. GROUP-LEVEL METRICS (HOST / STRAIN / SEX)
############################################################

group_metrics <- sample_metrics_full %>%
  group_by(Host, HostStrain, Sex) %>%
  summarise(
    n_samples          = n(),

    mean_count         = mean(DVG_count),
    mean_mean_length   = mean(mean_length, na.rm = TRUE),
    mean_p90_length    = mean(p90_length, na.rm = TRUE),
    mean_sd_length     = mean(sd_length, na.rm = TRUE),
    mean_IQR_length    = mean(IQR_length, na.rm = TRUE),

    mean_support       = mean(total_support),
    mean_sum_rpht      = mean(sum_rpht, na.rm = TRUE),

    mean_shannon_weighted = mean(shannon_weighted, na.rm = TRUE),
    mean_top1_fraction    = mean(top1_fraction, na.rm = TRUE),

    mean_entropy       = mean(BP_entropy, na.rm = TRUE),
    mean_fraction_poly = mean(fraction_polymerase_species, na.rm = TRUE),
    mean_n_segments    = mean(n_segments_with_dvg, na.rm = TRUE),

    mean_polymerase_support    = mean(polymerase_support),
    mean_fraction_polymerase_reads = mean(fraction_polymerase_reads, na.rm = TRUE),

    mean_inframe_reads_total     = mean(inframe_reads_total),
    mean_frameshift_reads_total  = mean(frameshift_reads_total),
    mean_frac_inframe_reads_total= mean(frac_inframe_reads_total, na.rm = TRUE),

    mean_inframe_events_total    = mean(inframe_events_total),
    mean_frameshift_events_total = mean(frameshift_events_total),
    mean_frac_inframe_events_total = mean(frac_inframe_events_total, na.rm = TRUE),

    mean_viral_reads_est = mean(viral_reads_est, na.rm = TRUE),
    mean_norm_richness   = mean(normalized_richness_rpht, na.rm = TRUE),
    mean_norm_abundance  = mean(normalized_abundance_rpht, na.rm = TRUE),

    .groups = "drop"
  )

write.csv(group_metrics,
          "Outputs/dvg_group_metrics.csv",
          row.names = FALSE)

############################################################
## 10. PREP FOR KW / WILCOXON
############################################################

sample_metrics_full <- sample_metrics_full %>%
  mutate(
    Host       = factor(Host),
    HostStrain = factor(HostStrain),
    Sex        = factor(Sex)
  )

run_kw_pw_host <- function(df, var_name, label) {

  y <- df[[var_name]]

  f  <- as.formula(paste(var_name, "~ Host"))
  kw <- kruskal.test(f, data = df)

  ws <- wilcox.test(df[[var_name]] ~ df$HostStrain)
  wx <- wilcox.test(df[[var_name]] ~ df$Sex)

  log_to_stats("\n==============================")
  log_to_stats("Metric:", label)
  log_to_stats("==============================")
  log_to_stats("\n>>> Host (Kruskal–Wallis)")
  log_block(kw)
  log_to_stats("\nHostStrain (Wilcoxon):")
  log_block(ws)
  log_to_stats("\nSex (Wilcoxon):")
  log_block(wx)

  results_list[[label]] <<- tibble(
    Metric              = label,
    KW_Host             = format_p(kw$p.value),
    Wilcoxon_HostStrain = format_p(ws$p.value),
    Wilcoxon_Sex        = format_p(wx$p.value)
  )
}

############################################################
## 11. RUN STATS (OLD + NEW METRICS)
############################################################

# all samples
run_kw_pw_host(sample_metrics_full, "DVG_count",     "DVG_count")
run_kw_pw_host(sample_metrics_full, "total_support","total_support")
run_kw_pw_host(sample_metrics_full, "sum_rpht",     "sum_rpht")

# new read-based polymerase fraction (all samples)
run_kw_pw_host(sample_metrics_full, "fraction_polymerase_reads", "fraction_polymerase_reads")

# DVG positive only
dvg_pos <- sample_metrics_full %>% filter(DVG_count > 0)

run_kw_pw_host(dvg_pos, "mean_length",   "mean_length")
run_kw_pw_host(dvg_pos, "sd_length",     "sd_length")
run_kw_pw_host(dvg_pos, "IQR_length",    "IQR_length")
run_kw_pw_host(dvg_pos, "BP_entropy",    "BP_entropy")
run_kw_pw_host(dvg_pos, "shannon_weighted", "shannon_weighted")
run_kw_pw_host(dvg_pos, "top1_fraction", "top1_fraction")
run_kw_pw_host(dvg_pos, "fraction_polymerase_species", "fraction_polymerase_species")
run_kw_pw_host(dvg_pos, "n_segments_with_dvg", "n_segments_with_dvg")
run_kw_pw_host(dvg_pos, "normalized_richness_rpht",  "normalized_richness_rpht")
run_kw_pw_host(dvg_pos, "normalized_abundance_rpht", "normalized_abundance_rpht")

# new frame-related metrics (DVG+ only)
run_kw_pw_host(dvg_pos, "frac_inframe_reads_total",   "frac_inframe_reads_total")
run_kw_pw_host(dvg_pos, "frac_inframe_events_total",  "frac_inframe_events_total")

results_df <- bind_rows(results_list)

write.csv(results_df,
          "Outputs/dvg_stats_summary_pvalues.csv",
          row.names = FALSE)

############################################################
## 12. DVG TYPE METRICS (UNCHANGED)
############################################################

dvg_filt_types <- dvg %>%
  filter(
    !is.na(DVG_type),
    !is.na(SampleID),
    mean_read_counts >= min_support
  )

dvg_type_counts <- dvg_filt_types %>%
  group_by(SampleID, Host, HostStrain, Sex, DVG_type) %>%
  summarise(n = n(), .groups="drop") %>%
  group_by(SampleID, Host, HostStrain, Sex) %>%
  tidyr::complete(DVG_type, fill=list(n=0)) %>%
  ungroup()

dvg_type_group_summary <- dvg_type_counts %>%
  group_by(DVG_type, Host, HostStrain, Sex) %>%
  summarise(
    mean   = mean(n),
    median = median(n),
    p90    = quantile(n,0.9),
    sd     = sd(n),
    .groups="drop"
  )

if (!dir.exists("Outputs/DVG_Types")) dir.create("Outputs/DVG_Types")
write.csv(dvg_type_group_summary,
          "Outputs/DVG_Types/dvg_type_group_summary.csv",
          row.names=FALSE)

```


_________________________
#### QUICK CORRELATIONS.
________________________

```{r}
# ============================================================
# CORRELATIONS WITH TCID50 AND COPY NUMBER
# ============================================================

corr_df <- sample_metrics_full %>%
  mutate(
    TCID50      = suppressWarnings(as.numeric(TCID50)),
    Copy.Number = suppressWarnings(as.numeric(Copy.Number))
  )

metric_list <- c(
  "DVG_count",
  "mean_length",
  "sd_length",
  "IQR_length",
  "total_support",
  "sum_rpht",
  "BP_entropy",
  "shannon_weighted",
  "top1_fraction",
  "fraction_polymerase_species",
  "fraction_polymerase_reads",
  "n_segments_with_dvg",
  "normalized_richness_rpht",
  "normalized_abundance_rpht"
)

# Convert ALL metric columns in metric_list to numeric
for (m in metric_list) {
  if (m %in% names(corr_df)) {
    corr_df[[m]] <- suppressWarnings(as.numeric(corr_df[[m]]))
  }
}


corr_fun <- function(x, y) {
  ok <- complete.cases(x, y)
  if (sum(ok) < 3) {
    tibble(correlation = NA_real_, p_value = NA_real_)
  } else {
    ct <- cor.test(x[ok], y[ok], method = "spearman", exact = FALSE)
    tibble(correlation = ct$estimate, p_value = ct$p.value)
  }
}



cor_results <- tibble(Measure = metric_list) %>%
  rowwise() %>%
  mutate(
    TCID50_corr = list(corr_fun(corr_df[[Measure]], corr_df$TCID50)),
    Copy_corr   = list(corr_fun(corr_df[[Measure]], corr_df$Copy.Number))
  ) %>%
  unnest_wider(TCID50_corr, names_sep = "_TCID50_") %>%
  unnest_wider(Copy_corr,   names_sep = "_Copy_")

print(cor_results)

write.csv(
  cor_results,
  "Outputs/dvg_correlations_TCID50_CopyNumber.csv",
  row.names = FALSE
)


```

AND WE PLOT.

```{r}
############################################################
#                     PLOTTING SECTION
############################################################

plot_theme <- theme_minimal() +
  theme(
    axis.title.x = element_text(size = 20, face = "bold", color = "black",
                                margin = margin(t = 15)),
    axis.title.y = element_text(size = 20, face = "bold", color = "black",
                                margin = margin(r = 15)),
    axis.text = element_text(size = 16, color = "black"),
    axis.line = element_line(size = 1, color = "black"),
    axis.ticks = element_line(size = 1),
    axis.ticks.length = unit(6, "pt"),
    panel.grid = element_blank(),
    legend.position = "none"     # clean: remove all legends
  )

cols <- c(
  BALBF = "#f94040",
  BALBM = "#0000FF",
  BL6F  = "#C5944E",
  BL6M  = "#0F99B2"
)

shapes <- c(
  BALBF = 1,
  BALBM = 22,
  BL6F  = 16,
  BL6M  = 15
)

# ============================================================
# Mean ± SE helper
# ============================================================

mean_se_df <- function(x) {
  m  <- mean(x, na.rm = TRUE)
  se <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
  data.frame(y = m, ymin = m - se, ymax = m + se)
}

# ============================================================
# Unified plotting function (Option 1)
# ============================================================

make_metric_plot <- function(df, metric, ylab, logscale = FALSE) {

  base <- ggplot(df, aes(x = Host, y = .data[[metric]], 
                         color = Host, shape = Host)) +

    # Mean horizontal line
    stat_summary(
      fun = mean,
      geom = "point",
      shape = 95,        # thick horizontal line
      size = 12,
      color = "black",
      stroke = 1
    ) +

    # SE whiskers
    stat_summary(
      fun.data = mean_se_df,
      geom = "errorbar",
      width = 0.15,
      linewidth = 0.6,
      color = "black"
    ) +

    # Individual points
    geom_jitter(size = 3.8, width = 0.20, alpha = 0.95, stroke = 1) +

    scale_color_manual(values = cols) +
    scale_shape_manual(values = shapes) +

    labs(x = "", y = ylab) +
    plot_theme

  if (logscale) {
    base <- base + scale_y_log10()
  }

  base
}

# ============================================================
# Output folder
# ============================================================

if (!dir.exists("Images/Metrics"))
  dir.create("Images/Metrics", recursive = TRUE)

# ============================================================
# Updated full list of metrics to plot
# ============================================================

metrics_to_plot <- list(

  # richness / abundance
  list(var = "DVG_count",                 label = "Unique DVGs",                         log = FALSE),
  list(var = "total_support",             label = "Total Support (reads)",               log = TRUE),
  list(var = "sum_rpht",                  label = "Sum RPHT",                            log = TRUE),

  # normalized metrics
  list(var = "normalized_richness_rpht",  label = "Normalized Richness (per 1e5)",       log = FALSE),
  list(var = "normalized_abundance_rpht", label = "Normalized Abundance (per 1e5)",      log = TRUE),

  # length distribution
  list(var = "mean_length",               label = "Mean Deletion Length",                     log = FALSE),
  list(var = "median_length",             label = "Median Deletion Length",                   log = FALSE),
  list(var = "p90_length",                label = "P90 Length",                          log = FALSE),
  list(var = "sd_length",                 label = "SD Length",                           log = FALSE),
  list(var = "IQR_length",                label = "IQR Length",                          log = FALSE),

  # population structure
  list(var = "BP_entropy",                label = "Breakpoint Entropy",                  log = FALSE),
  list(var = "shannon_weighted",          label = "Shannon (weighted by reads)",         log = FALSE),
  list(var = "top1_fraction",             label = "Top DVG Fraction",                    log = FALSE),

  # segment structure
  list(var = "fraction_polymerase_species",       label = "Fraction Polymerase-derived Species",         log = FALSE),
  list(var = "fraction_polymerase_reads",       label = "Fraction Polymerase-derived Reads",         log = FALSE),
  list(var = "n_segments_with_dvg",       label = "Number of Segments with DVGs",        log = FALSE),

  # viral load estimate
  list(var = "viral_reads_est",           label = "Estimated Viral Reads",               log = TRUE)
)

# ============================================================
# Generate each plot automatically
# ============================================================

for (m in metrics_to_plot) {

  df_sub <- sample_metrics_full %>% filter(!is.na(.data[[m$var]]))

  p <- make_metric_plot(
    df = df_sub,
    metric = m$var,
    ylab = m$label,
    logscale = m$log
  )

  ggsave(
    paste0("Images/Metrics/", m$var, ".png"),
    p, width = 7, height = 5, dpi = 300, bg = "white"
  )
  ggsave(
    paste0("Images/Metrics/", m$var, ".svg"),
    p, width = 7, height = 5, dpi = 300, bg = "white"
  )

  print(p)
}


```

ECDF Mean length

```{r}
# ============================================================
# Prism-like theme (your version)
# ============================================================

plot_theme <- theme_minimal() +
  theme(
    axis.title.x = element_text(size = 24, face = "bold", color = "black",
                                margin = margin(t = 15)),
    axis.title.y = element_text(size = 24, face = "bold", color = "black",
                                margin = margin(r = 15)),
    axis.text = element_text(size = 18, color = "black"),
    axis.line = element_line(size = 1, color = "black"),
    axis.ticks = element_line(size = 1),
    axis.ticks.length = unit(8, "pt"),
    panel.grid = element_blank(),
    legend.position = "none",
    legend.text = element_text(size = 18),
    legend.title = element_blank()
  )

# Colors
cols <- c(
  BALBF = "#f94040",
  BALBM = "#0000FF",
  BL6F  = "#C5944E",
  BL6M  = "#0F99B2"
)

# Line styles:
# BALB groups dashed, BL6 groups solid
line_types <- c(
  BALBF = "dashed",
  BALBM = "dashed",
  BL6F  = "solid",
  BL6M  = "solid"
)

# ============================================================
# ECDF PLOT with custom x-axis limits
# ============================================================

# Adjust these as needed:
x_min <- 0
x_max <- 2300   # <-- change this to whatever you want

p_ecdf <- ggplot(dvg_filt, aes(x = length_dvg, color = Host, linetype = Host)) +
  stat_ecdf(geom = "step", linewidth = 1) +

  scale_color_manual(values = cols) +
  scale_linetype_manual(values = line_types) +

  scale_x_continuous(limits = c(x_min, x_max)) +

  labs(
    x = "DVG Length",
    y = "ECDF"
  ) +

  plot_theme +
  theme(
    legend.title = element_blank()
  )

p_ecdf

# Save
ggsave("Images/Metrics/ECDF_length.png",
       p_ecdf, width = 8, height = 5, dpi = 300, bg = "white")

ggsave("Images/Metrics/ECDF_length.svg",
       p_ecdf, width = 8, height = 5, dpi = 300, bg = "white")

```


____________________
##### SEGMENT PLOTS
____________________
By support. Change alpha of lines before/after/between ticks to show either the inside of deletion or the outside (genome before and after break)

```{r}
## ============================================================
## DVG line plots: BP / RI on x, per-DVG support on y
## ============================================================

# List of segments and their approximate plotting x limits
segs  <- c("PB1", "PB2", "PA", "HA", "NP", "NR", "M", "NS")
xlims <- c(2285, 2284, 2183, 1717, 1506, 1426,  985,  847)

colors <- c(
  "BALBF" = "#F94040",
  "BALBM" = "#0000FF",
  "BL6F"  = "#C5944E",
  "BL6M"  = "#0F99B2"
)

# Segment genomic coordinates (start, end)
segment_ends <- list(
  PB1 = c(1, 2341),
  PB2 = c(1, 2341),
  PA  = c(1, 2233),
  HA  = c(1, 1778),
  NP  = c(1, 1565),
  NR  = c(1, 1433),
  M   = c(1, 1027),
  NS  = c(1,  890)
)

# Loop over segments
for (i in seq_along(segs)) {

  segment  <- segs[i]
  xlim_max <- xlims[i]

  # Use already filtered DVGs (by support etc.)
  df1 <- dvg_filt %>%
    filter(Segment == segment)

  if (nrow(df1) == 0) {
    message("No DVGs for segment ", segment, ", skipping.")
    next
  }

  # y is per-DVG support (mean_read_counts from DVGfinder)
  # set a reasonable y-range per segment
  y_max <- max(df1$mean_read_counts, na.rm = TRUE)
  if (!is.finite(y_max) || y_max <= 0) {
    message("Non-finite y_max for segment ", segment, ", skipping.")
    next
  }

  # vertical half-height of the start / end ticks, scaled to the y-range
  delheight <- y_max * 0.02  # 2 percent of max support
  delwidth  <- 0.7           # line width for ticks

  # dummy data to ensure all host colors are in scale (even if absent for this segment)
  dummy_data <- expand.grid(
    SampleID   = "dummy",
    Sex        = "X",
    HostStrain = "Dummy",
    Host       = names(colors),
    Treatment  = "Dummy",
    Type       = "Dummy",
    Line       = "Dummy",
    Virulence  = 0,
    TCID50     = 0,
    Copy.Number = 0,
    npoly      = 0,
    Nei_D      = 0,
    geneDiv    = 0,
    Segment    = segment,
    BP         = 1,
    RI         = 2,
    mean_read_counts = 0
  )

  df1_plot <- bind_rows(df1, dummy_data)

  seg_start <- segment_ends[[segment]][1]
  seg_end   <- segment_ends[[segment]][2]

  p <- ggplot(df1_plot, aes(color = Host)) +

    # Thin line from segment start to BP
    geom_segment(
      aes(x = seg_start, xend = BP,
          y = mean_read_counts, yend = mean_read_counts),
      linewidth = 0.2,
      linetype  = "dotted",
      alpha = 1
    ) +

    # Tick at BP
    geom_segment(
      aes(x = BP, xend = BP,
          y = mean_read_counts - delheight,
          yend = mean_read_counts + delheight),
      linewidth = delwidth
    ) +

    # Horizontal deletion line between BP and RI
    geom_segment(
      aes(x = BP, xend = RI,
          y = mean_read_counts, yend = mean_read_counts),
      linewidth = 0.2,
      linetype  = "dotted",
      alpha = 0
    ) +

    # Tick at RI
    geom_segment(
      aes(x = RI, xend = RI,
          y = mean_read_counts - delheight,
          yend = mean_read_counts + delheight),
      linewidth = delwidth
    ) +

    # Thin line from RI to segment end
    geom_segment(
      aes(x = RI, xend = seg_end,
          y = mean_read_counts, yend = mean_read_counts),
      linewidth = 0.2,
      linetype  = "dotted",
      alpha = 1
    ) +

    scale_color_manual(values = colors, drop = FALSE) +
    coord_cartesian(
      xlim = c(seg_start, seg_end),
      expand = FALSE
    ) +
    scale_x_continuous(breaks = seq(0, seg_end, by = 250)) +
    # let ggplot choose y breaks; you can also set them manually if you prefer
    scale_y_continuous() +
    theme_minimal() +
    theme(
      plot.margin      = unit(c(1, 1, 2, 1), "lines"),
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background  = element_rect(fill = "white", colour = NA),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position  = "none",
      axis.text        = element_text(size = 16, color = "black"),
      axis.title       = element_text(size = 20, face = "bold", color = "black"),
      axis.title.x     = element_text(vjust = -1),
      axis.title.y     = element_text(vjust = +2),
      axis.ticks       = element_line(colour = "black", linewidth = 1),
      axis.ticks.length = unit(.25, "cm"),
      axis.line        = element_line(color = "black", linewidth = 1)
    ) +
    labs(
      x = "Genomic position",
      y = "DVG read support"
    )

  ggsave(
    paste0("Images/DVG Segment Plots - Support/DVGlines_", segment, "_genome.png"),
    p, width = 8, height = 5, dpi = 300
  )
  ggsave(
    paste0("Images/DVG Segment Plots - Support/DVGlines_", segment, "_genome.svg"),
    p, width = 8, height = 5, dpi = 300, device = "svg"
  )
}

```



Support, combined segments
```{r}
###############################################################
## 1. SETTINGS
###############################################################

library(dplyr)
library(ggplot2)
library(patchwork)
library(scales)

segment_order <- c("PB2","PB1","PA","HA","NP","NR","M","NS")

segment_ends <- list(
  PB2 = c(1, 2341),
  PB1 = c(1, 2341),
  PA  = c(1, 2233),
  HA  = c(1, 1778),
  NP  = c(1, 1565),
  NR  = c(1, 1433),
  M   = c(1, 1027),
  NS  = c(1,  890)
)

host_order <- c("BALBF","BALBM","BL6F","BL6M")

colors <- c(
  "BALBF" = "#F94040",
  "BALBM" = "#0000FF",
  "BL6F"  = "#C5944E",
  "BL6M"  = "#0F99B2"
)


###############################################################
## 2. PREP DATA
###############################################################

stopifnot(all(c("mean_read_counts","BP","RI","Segment","SampleID","Host") %in% names(dvg_filt)))

dvg_sup <- dvg_filt %>%
  filter(
    !is.na(Segment),
    Segment %in% segment_order,
    !is.na(BP),
    !is.na(RI)
  ) %>%
  mutate(
    Segment = factor(Segment, levels = segment_order),
    Host    = factor(Host,    levels = host_order)
  ) %>%
  group_by(SampleID, Segment) %>%
  mutate(
    support_segment = mean_read_counts
  ) %>%
  ungroup()

# global max (for y-axis limit)
y_max_support <- max(dvg_sup$support_segment, na.rm = TRUE)
y_max_support <- ceiling(y_max_support * 1.2)

# tick height factor for log axis
tick_factor <- 1.1


###############################################################
## 3. BUILD MASTER DF FOR COMBINED PLOT
###############################################################

full_plot_df <- bind_rows(lapply(segment_order, function(seg) {
  df <- dvg_sup %>% filter(Segment == seg)
  if (nrow(df)==0) return(NULL)
  df$seg_start <- segment_ends[[seg]][1]
  df$seg_end   <- segment_ends[[seg]][2]
  df
}))

full_plot_df$Segment <- factor(full_plot_df$Segment, levels = segment_order)


###############################################################
## 4. FUNCTION TO PLOT ONE SEGMENT (COMBINED PANEL)
###############################################################

plot_seg_support <- function(df_seg, seg_name, first=FALSE){

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]

  ggplot(df_seg, aes(color = Host)) +

    # genome left of BP
    geom_segment(aes(
      x = seg_start, xend = BP,
      y = support_segment, yend = support_segment),
      linewidth = 0.2, linetype = "dotted", alpha = 0
    ) +

    # BP tick (log-scale)
    geom_segment(aes(
      x = BP, xend = BP,
      y = support_segment / tick_factor,
      yend = support_segment * tick_factor),
      linewidth = 1, alpha = 0.8
    ) +

    # deletion
    geom_segment(aes(
      x = BP, xend = RI,
      y = support_segment, yend = support_segment),
      linewidth = 0.2, alpha = 0.8, linetype = "dotted"
    ) +

    # RI tick (log-scale)
    geom_segment(aes(
      x = RI, xend = RI,
      y = support_segment / tick_factor,
      yend = support_segment * tick_factor),
      linewidth = 1, alpha = 0.8
    ) +

    # genome right of RI
    geom_segment(aes(
      x = RI, xend = seg_end,
      y = support_segment, yend = support_segment),
      linewidth = 0.2, linetype = "dotted", alpha = 0
    ) +

    scale_color_manual(values = colors) +

    # X axis
    scale_x_continuous(
      limits = c(seg_start, seg_end),
      breaks = seq(0, seg_end, by = 500),
      expand = c(0,0)
    ) +

    # LOG Y
    scale_y_log10(
      limits = c(5, y_max_support),
      breaks = c(10, 100, 1000, 10000),
      labels = c("10","100","1000","10000"),
      expand = c(0, 0)
    ) +

    labs(
      x = seg_name,
      y = if(first) "DVG read support" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid = element_blank(),

      legend.position = "none",

      axis.text.x  = element_text(size = 10, color = "black"),
      axis.title.x = element_text(size = 12, face = "bold"),

      axis.text.y  = if(first) element_text(size = 10) else element_blank(),
      axis.title.y = if(first) element_text(size = 12, face = "bold") else element_blank(),
      axis.ticks.y = if(first) element_line(color="black") else element_blank(),
      axis.line.y  = if(first) element_line(color="black") else element_blank(),

      axis.line.x  = element_line(color="black"),
      axis.ticks.x = element_line(color="black"),
      axis.ticks.length = unit(3, "pt")
    )
}


###############################################################
## 5. CREATE PANELS FOR ALL SEGMENTS AND COMBINE
###############################################################

plots <- lapply(seq_along(segment_order), function(i){
  seg <- segment_order[i]
  df_seg <- full_plot_df %>% filter(Segment==seg)
  plot_seg_support(df_seg, seg, first=(i==1))
})

segment_lengths <- sapply(segment_ends, function(x) x[2] - x[1])
relative_widths <- segment_lengths[segment_order] / sum(segment_lengths)

combined_sup <- wrap_plots(plots, nrow=1, widths=relative_widths)


###############################################################
## 6. SAVE
###############################################################

if (!dir.exists("Images/DVG Segment Plots - combined"))
  dir.create("Images/DVG Segment Plots - combined", recursive = TRUE)

ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_support_log.png",
       combined_sup, width=16, height=8, dpi=300)

ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_support_log.svg",
       combined_sup, width=16, height=8, dpi=300)


```


By fraction of support
```{r}
## ============================================================
## DVG line plots: BP / RI on x, fraction of DVG reads per sample
## ============================================================

# List of segments and their approximate plotting x limits
segs  <- c("PB1", "PB2", "PA", "HA", "NP", "NR", "M", "NS")
xlims <- c(2285, 2284, 2183, 1717, 1506, 1426,  985,  847)

colors <- c(
  "BALBF" = "#F94040",
  "BALBM" = "#0000FF",
  "BL6F"  = "#C5944E",
  "BL6M"  = "#0F99B2"
)

# Segment genomic coordinates (start, end)
segment_ends <- list(
  PB1 = c(1, 2341),
  PB2 = c(1, 2341),
  PA  = c(1, 2233),
  HA  = c(1, 1778),
  NP  = c(1, 1565),
  NR  = c(1, 1433),
  M   = c(1, 1027),
  NS  = c(1,  890)
)


dvg_filt <- dvg %>%
  filter(
    !is.na(BP),
    !is.na(RI),
    length_dvg > 0,
    mean_read_counts >= min_support
  )


# Compute total DVG support per sample
sample_total_support <- dvg_filt %>%
  group_by(SampleID) %>%
  summarise(total_support = sum(mean_read_counts), .groups="drop")

# Add fractional support to dvg_filt
dvg_filt <- dvg_filt %>%
  left_join(sample_total_support, by = "SampleID") %>%
  mutate(
    frac_support = mean_read_counts / total_support
  )

# dvg_filt now contains frac_support

for (i in seq_along(segs)) {

  segment  <- segs[i]
  xlim_max <- xlims[i]

  df1 <- dvg_filt %>% filter(Segment == segment)

  if (nrow(df1) == 0) {
    message("No DVGs for segment ", segment, ", skipping.")
    next
  }

  if (!is.finite(y_max) || y_max <= 0) next

  delheight <- 0.01
  delwidth  <- 0.7

  dummy_data <- expand.grid(
    SampleID   = "dummy",
    Sex        = "X",
    HostStrain = "Dummy",
    Host       = names(colors),
    Treatment  = "Dummy",
    Type       = "Dummy",
    Line       = "Dummy",
    Virulence  = 0,
    TCID50     = 0,
    Copy.Number = 0,
    npoly      = 0,
    Nei_D      = 0,
    geneDiv    = 0,
    Segment    = segment,
    BP         = 1,
    RI         = 2,
    frac_support = 0
  )

  df1_plot <- bind_rows(df1, dummy_data)

  seg_start <- segment_ends[[segment]][1]
  seg_end   <- segment_ends[[segment]][2]

  p <- ggplot(df1_plot, aes(color = Host)) +

    # Segment start → BP
    geom_segment(aes(x=seg_start, xend=BP, 
                     y=frac_support, yend=frac_support),
                 linewidth=0.2, linetype="dotted") +

    # Tick at BP
    geom_segment(aes(x=BP, xend=BP,
                     y=frac_support - delheight,
                     yend=frac_support + delheight),
                 linewidth=delwidth) +

    # BP → RI (hidden line)
    geom_segment(aes(x=BP, xend=RI,
                     y=frac_support, yend=frac_support),
                 linewidth=0.2, alpha=0) +

    # Tick at RI
    geom_segment(aes(x=RI, xend=RI,
                     y=frac_support - delheight,
                     yend=frac_support + delheight),
                 linewidth=delwidth) +

    # RI → segment end
    geom_segment(aes(x=RI, xend=seg_end,
                     y=frac_support, yend=frac_support),
                 linewidth=0.2, linetype="dotted") +
    
    annotate(
    "text",
    x = seg_start + (seg_end - seg_start) / 2,   # centered horizontally
    y = 0.48,                                     # just below top of axis
    label = segment,
    size = 7,
    fontface = "bold",
    color = "black"
  ) +

    scale_color_manual(values=colors, drop=FALSE) +

    coord_cartesian(
      xlim = c(seg_start, seg_end),
      expand = FALSE
    ) +

    scale_x_continuous(breaks = seq(0, seg_end, by=250)) +
    scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, 0.5, by = 0.1), expand = c(0, 0))+

    theme_minimal() +
    theme(
      plot.margin      = unit(c(1, 1, 2, 1), "lines"),
      panel.background = element_rect(fill="white", colour=NA),
      plot.background  = element_rect(fill="white", colour=NA),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position  = "none",
      axis.text        = element_text(size=16, color="black"),
      axis.title       = element_text(size=20, face="bold", color="black"),
      axis.title.x     = element_text(vjust=-1),
      axis.title.y     = element_text(vjust=+2),
      axis.ticks       = element_line(color="black", linewidth=1),
      axis.ticks.length = unit(.25, "cm"),
      axis.line        = element_line(color="black", linewidth=1)
    ) +
    labs(
      x = "Genomic position",
      y = "Fraction of DVG reads"
    )

  ggsave(paste0("Images/DVG Segment Plots - Fraction/DVGlines_", segment, ".png"),
         p, width=8, height=5, dpi=300)

  ggsave(paste0("Images/DVG Segment Plots - Fraction/DVGlines_", segment, ".svg"),
         p, width=8, height=5, dpi=300, device="svg")
}

```

All segments in one graph

```{r}
###########################################
# All segments in one graph
###########################################
segment_order <- c("PB2", "PB1", "PA", "HA", "NP", "NR", "M", "NS")

segment_ends <- list(
  PB1 = c(1, 2341),
  PB2 = c(1, 2341),
  PA  = c(1, 2233),
  HA  = c(1, 1778),
  NP  = c(1, 1565),
  NR  = c(1, 1433),
  M   = c(1, 1027),
  NS  = c(1,  890)
)

colors <- c(
  "BALBF"="#F94040",
  "BALBM"="#0000FF",
  "BL6F" ="#C5944E",
  "BL6M" ="#0F99B2"
)

###########################################
# Build full plotting df for ALL segments
###########################################

tick_height <- 0.01
tick_width  <- 0.7

full_plot_df <- lapply(segment_order, function(seg) {
  seg_start <- segment_ends[[seg]][1]
  seg_end   <- segment_ends[[seg]][2]

  df <- dvg_filt %>% filter(Segment == seg)
  if (nrow(df) == 0) return(NULL)

  df$seg_start <- seg_start
  df$seg_end   <- seg_end
  df
}) %>%
  bind_rows()

###########################################
# Enforce Segment ordering
###########################################
full_plot_df$Segment <- factor(full_plot_df$Segment,
                               levels = segment_order)

###########################################
# PLOTTING
###########################################

p <- ggplot(full_plot_df, aes(color = Host)) +

  geom_segment(aes(x = seg_start, xend = BP,
                   y = frac_support, yend = frac_support),
               linewidth = 0.2, linetype = "dotted") +

  geom_segment(aes(x = BP, xend = BP,
                   y = frac_support - tick_height,
                   yend = frac_support + tick_height),
               linewidth = tick_width) +

  geom_segment(aes(x = BP, xend = RI,
                   y = frac_support, yend = frac_support),
               alpha = 0) +

  geom_segment(aes(x = RI, xend = RI,
                   y = frac_support - tick_height,
                   yend = frac_support + tick_height),
               linewidth = tick_width) +

  geom_segment(aes(x = RI, xend = seg_end,
                   y = frac_support, yend = frac_support),
               linewidth = 0.2, linetype = "dotted") +

  scale_color_manual(values = colors) +

  facet_wrap(~Segment, nrow = 1, scales = "free_x") +

  scale_y_continuous(
    limits = c(0, 0.5),
    breaks = seq(0, 0.5, by = 0.1),
    expand = c(0, 0)
  ) +

  labs(
    x = "Genomic position",
    y = "Fractional of DVG reads"
  ) +

  theme_minimal() +
  theme(
    strip.text = element_text(size = 18, face = "bold"),
    axis.text  = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 18, face = "bold"),
    panel.grid = element_blank(),
    legend.position = "none",
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black")
  )

p

ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_facet.png",
       p, width = 18, height = 6, dpi = 300)

ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_facet.svg",
       p, width = 18, height = 6, dpi = 300)
```


Combined segments - genome painted
```{r}
############################################
# Segment ordering (largest → smallest)
############################################
segment_order <- c("PB2","PB1","PA","HA","NP","NR","M","NS")

segment_ends <- list(
  PB2 = c(1,2341),
  PB1 = c(1,2341),
  PA  = c(1,2233),
  HA  = c(1,1778),
  NP  = c(1,1565),
  NR  = c(1,1433),
  M   = c(1,1027),
  NS  = c(1, 890)
)

colors <- c(
  "BALBF"="#F94040",
  "BALBM"="#0000FF",
  "BL6F" ="#C5944E",
  "BL6M" ="#0F99B2"
)

tick_height <- 0.01
tick_width  <- 0.7


#######################################################
# Function to draw ONE segment panel (whole genome fraction)
#######################################################
plot_segment <- function(df_seg, seg_name, is_first=FALSE){

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]

  sec_breaks <- seq(250, seg_end - 250, by = 500)

  ggplot(df_seg, aes(color = Host)) +

    # start → BP
    geom_segment(aes(
      x = seg_start, xend = BP,
      y = frac_support, yend = frac_support
    ),
    linewidth = 0.4, linetype = "dotted", alpha =0.5) +

    # BP tick
    geom_segment(aes(
      x = BP, xend = BP,
      y = frac_support - tick_height,
      yend = frac_support + tick_height
    ),
    linewidth = tick_width, lineend = "butt",alpha=0.6) +

    # BP → RI (invisible)
    geom_segment(aes(
      x = BP, xend = RI,
      y = frac_support, yend = frac_support
    ),
    alpha = 0) +

    # RI tick
    geom_segment(aes(
      x = RI, xend = RI,
      y = frac_support - tick_height,
      yend = frac_support + tick_height
    ),
    linewidth = tick_width, lineend = "butt",alpha=0.6) +

    # RI → end
    geom_segment(aes(
      x = RI, xend = seg_end,
      y = frac_support, yend = frac_support
    ),
    linewidth = 0.4, linetype = "dotted", alpha=0.5) +

    scale_color_manual(values = colors) +

    scale_x_continuous(
      limits = c(seg_start, seg_end),
      breaks = seq(0, seg_end, 500),
      minor_breaks = sec_breaks,
      expand = c(0,0)
    ) +

    scale_y_continuous(
      limits = c(0, 0.5),
      breaks = seq(0, 0.5, 0.1),
      expand = c(0, 0)
    ) +

    labs(
      x = seg_name,
      y = if (is_first) "Fraction of DVG reads" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",

      axis.text.x  = element_text(size = 12, color = "black"),
      axis.title.x = element_text(size = 14, face = "bold"),

      # y-axis ONLY for PB2
      axis.line.y  = if(is_first) element_line(color="black", linewidth=1) else element_blank(),
      axis.ticks.y = if(is_first) element_line(color="black") else element_blank(),
      axis.text.y  = if(is_first) element_text(size=12, color="black") else element_blank(),
      axis.title.y = if(is_first) element_text(size=14, face="bold") else element_blank(),

      # x-axis always visible
      axis.line.x  = element_line(color="black", linewidth=1),
      axis.ticks.x = element_line(color="black"),
      axis.ticks.length = unit(4, "pt")
    )
}


#######################################################
# Build master dataframe for all segments
#######################################################
full_plot_df <- bind_rows(lapply(segment_order, function(seg) {

  seg_start <- segment_ends[[seg]][1]
  seg_end   <- segment_ends[[seg]][2]

  df <- dvg_filt %>% filter(Segment == seg)
  if(nrow(df) == 0) return(NULL)

  df$seg_start <- seg_start
  df$seg_end   <- seg_end

  df
}))


#######################################################
# Create one plot per segment
#######################################################
plots <- lapply(seq_along(segment_order), function(i){
  seg <- segment_order[i]
  df_seg <- full_plot_df %>% filter(Segment == seg)
  plot_segment(df_seg, seg, is_first = (i == 1))
})


#######################################################
# Combine with proportional widths
#######################################################
segment_lengths <- sapply(segment_ends, function(x) x[2] - x[1])
relative_widths <- segment_lengths[segment_order] / sum(segment_lengths)

combined <- wrap_plots(plots, nrow = 1, widths = relative_widths)


#######################################################
# Save output
#######################################################
ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_final.png",
       combined, width = 18, height = 6, dpi = 300)

ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_final.svg",
       combined, width = 18, height = 6, dpi = 300)

```


SAME AS ABOVE, deletion painted
```{r}
############################################
# Segment ordering (largest → smallest)
############################################
segment_order <- c("PB2","PB1","PA","HA","NP","NR","M","NS")

segment_ends <- list(
  PB2 = c(1,2341),
  PB1 = c(1,2341),
  PA  = c(1,2233),
  HA  = c(1,1778),
  NP  = c(1,1565),
  NR  = c(1,1433),
  M   = c(1,1027),
  NS  = c(1, 890)
)

colors <- c(
  "BALBF"="#F94040",
  "BALBM"="#0000FF",
  "BL6F" ="#C5944E",
  "BL6M" ="#0F99B2"
)

tick_height <- 0.01
tick_width  <- 0.7


#######################################################
# Function to draw ONE segment panel (whole genome fraction)
#######################################################
plot_segment <- function(df_seg, seg_name, is_first=FALSE){

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]

  sec_breaks <- seq(250, seg_end - 250, by = 500)

  ggplot(df_seg, aes(color = Host)) +

    # start → BP #(invisible)
    geom_segment(aes(
      x = seg_start, xend = BP,
      y = frac_support, yend = frac_support
    ),
    alpha =0) + 

    # BP tick
    geom_segment(aes(
      x = BP, xend = BP,
      y = frac_support - tick_height,
      yend = frac_support + tick_height
    ),
    linewidth = tick_width, lineend = "butt",alpha=0.7) +

    # BP → RI 
    geom_segment(aes(
      x = BP, xend = RI,
      y = frac_support, yend = frac_support
    ),
    linewidth = 0.4, linetype = "dotted", alpha = 0.5) +

    # RI tick
    geom_segment(aes(
      x = RI, xend = RI,
      y = frac_support - tick_height,
      yend = frac_support + tick_height
    ),
    linewidth = tick_width, lineend = "butt",alpha=0.7) +

    # RI → end #(invisible)
    geom_segment(aes(
      x = RI, xend = seg_end,
      y = frac_support, yend = frac_support
    ),
    alpha=0) +

    scale_color_manual(values = colors) +

    scale_x_continuous(
      limits = c(seg_start, seg_end),
      breaks = seq(0, seg_end, 500),
      minor_breaks = sec_breaks,
      expand = c(0,0)
    ) +

    scale_y_continuous(
      limits = c(0, 0.5),
      breaks = seq(0, 0.5, 0.1),
      expand = c(0, 0)
    ) +

    labs(
      x = seg_name,
      y = if (is_first) "Fraction of DVG reads" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",

      axis.text.x  = element_text(size = 12, color = "black"),
      axis.title.x = element_text(size = 14, face = "bold"),

      # y-axis ONLY for PB2
      axis.line.y  = if(is_first) element_line(color="black", linewidth=1) else element_blank(),
      axis.ticks.y = if(is_first) element_line(color="black") else element_blank(),
      axis.text.y  = if(is_first) element_text(size=12, color="black") else element_blank(),
      axis.title.y = if(is_first) element_text(size=14, face="bold") else element_blank(),

      # x-axis always visible
      axis.line.x  = element_line(color="black", linewidth=1),
      axis.ticks.x = element_line(color="black"),
      axis.ticks.length = unit(4, "pt")
    )
}


#######################################################
# Build master dataframe for all segments
#######################################################
full_plot_df <- bind_rows(lapply(segment_order, function(seg) {

  seg_start <- segment_ends[[seg]][1]
  seg_end   <- segment_ends[[seg]][2]

  df <- dvg_filt %>% filter(Segment == seg)
  if(nrow(df) == 0) return(NULL)

  df$seg_start <- seg_start
  df$seg_end   <- seg_end

  df
}))


#######################################################
# Create one plot per segment
#######################################################
plots <- lapply(seq_along(segment_order), function(i){
  seg <- segment_order[i]
  df_seg <- full_plot_df %>% filter(Segment == seg)
  plot_segment(df_seg, seg, is_first = (i == 1))
})


#######################################################
# Combine with proportional widths
#######################################################
segment_lengths <- sapply(segment_ends, function(x) x[2] - x[1])
relative_widths <- segment_lengths[segment_order] / sum(segment_lengths)

combined <- wrap_plots(plots, nrow = 1, widths = relative_widths)


#######################################################
# Save output
#######################################################
ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_delpainted.png",
       combined, width = 18, height = 6, dpi = 300)

ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_delpainted.svg",
       combined, width = 18, height = 6, dpi = 300)

```


```{r}
########################################################
## FRACTIONAL SUPPORT PLOTS — HOSTS SEPARATED       
########################################################

segment_order <- c("PB2","PB1","PA","HA","NP","NR","M","NS")

segment_ends <- list(
  PB2 = c(1,2341),
  PB1 = c(1,2341),
  PA  = c(1,2233),
  HA  = c(1,1778),
  NP  = c(1,1565),
  NR  = c(1,1433),
  M   = c(1,1027),
  NS  = c(1, 890)
)

colors <- c(
  "BALBF"="#F94040",
  "BALBM"="#0000FF",
  "BL6F" ="#C5944E",
  "BL6M" ="#0F99B2"
)

tick_height <- 0.01
tick_width  <- 0.7


########################################################
## FUNCTION: ONE SEGMENT PANEL
########################################################
plot_segment_host <- function(df_seg, seg_name, first_seg){

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]
  sec_breaks <- seq(250, seg_end - 250, by = 500)

  ggplot(df_seg, aes(color = Host)) +

    geom_segment(aes(x = seg_start, xend = BP,
                     y = frac_support, yend = frac_support),
                 linewidth = 0.2, linetype = "dotted") +

    geom_segment(aes(x = BP, xend = BP,
                     y = frac_support - tick_height,
                     yend = frac_support + tick_height),
                 linewidth = tick_width) +

    geom_segment(aes(x = BP, xend = RI,
                     y = frac_support, yend = frac_support),
                 alpha = 0) +

    geom_segment(aes(x = RI, xend = RI,
                     y = frac_support - tick_height,
                     yend = frac_support + tick_height),
                 linewidth = tick_width) +

    geom_segment(aes(x = RI, xend = seg_end,
                     y = frac_support, yend = frac_support),
                 linewidth = 0.2, linetype = "dotted") +

    scale_color_manual(values = colors) +

    scale_x_continuous(
      limits = c(seg_start, seg_end),
      breaks = seq(0, seg_end, 500),
      minor_breaks = sec_breaks,
      expand = c(0,0)
    ) +
    scale_y_continuous(
      limits = c(0,0.5),
      breaks = seq(0,0.5,0.1),
      expand = c(0,0)
    ) +

    labs(
      x = seg_name,
      y = if(first_seg) "Fraction DVG reads" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",

      axis.text.x  = element_text(size = 12, color = "black"),
      axis.title.x = element_text(size = 13, face = "bold"),

      # only PB2 gets y axis
      axis.line.y  = if(first_seg) element_line(color="black") else element_blank(),
      axis.ticks.y = if(first_seg) element_line(color="black") else element_blank(),
      axis.text.y  = if(first_seg) element_text(size=12, color="black") else element_blank(),
      axis.title.y = if(first_seg) element_text(size=14, face="bold") else element_blank(),

      axis.line.x  = element_line(color="black"),
      axis.ticks.x = element_line(color="black"),
      axis.ticks.length = unit(4, "pt")
    )
}


########################################################
## FUNCTION: FULL HOST PANEL (8 SEGMENTS)
########################################################

plot_one_host <- function(df_host){

  seg_plots <- lapply(seq_along(segment_order), function(i){

    seg <- segment_order[i]
    df_seg <- df_host %>% filter(Segment == seg)

    if(nrow(df_seg)==0){
      return(ggplot() + theme_void())
    }

    plot_segment_host(df_seg, seg, first_seg = (i==1))
  })

  segment_lengths <- sapply(segment_ends, function(x) x[2]-x[1])
  relative_widths <- segment_lengths[segment_order] / sum(segment_lengths)

  wrap_plots(seg_plots, nrow = 1, widths = relative_widths)
}


########################################################
## PREP DATA: full_plot_df MUST EXIST
########################################################
## (Your earlier pipeline already created dvg_filt
##  and full_plot_df; just make sure frac_support exists.)

## For safety:
stopifnot("frac_support" %in% colnames(full_plot_df))


########################################################
## GENERATE FOUR HOST PANELS
########################################################

hosts <- c("BALBF","BALBM","BL6F","BL6M")

p_list <- list()

for(h in hosts){

  df_host <- full_plot_df %>% filter(Host == h)

  if(nrow(df_host)==0){
    message("Skipping ", h)
    p_list[[h]] <- ggplot() + theme_void()
    next
  }

  p_host <- plot_one_host(df_host)

  # store in list
  p_list[[h]] <- p_host

  # save individual
  ggsave(
    paste0("Images/DVG Segment Plots - byHost/DVG_", h, "_fractional.png"),
    p_host, width = 18, height = 5, dpi = 300
  )
    ggsave(
    paste0("Images/DVG Segment Plots - byHost/DVG_", h, "_fractional.svg"),
    p_host, width = 18, height = 5, dpi = 300
  )
}

# Assign to variables for final stacking
p_BALBF <- p_list[["BALBF"]]
p_BALBM <- p_list[["BALBM"]]
p_BL6F  <- p_list[["BL6F"]]
p_BL6M  <- p_list[["BL6M"]]


########################################################
## STACKED FIGURE (ONLY BL6M HAS LEGEND)
########################################################

stacked <- 
  (p_BALBF + theme(legend.position="none")) /
  (p_BALBM + theme(legend.position="none")) /
  (p_BL6F  + theme(legend.position="none")) /
  (p_BL6M + theme(legend.position="none")) +
  plot_layout(heights = c(1,1,1,1.2))

ggsave("Images/DVG Segment Plots - combined/DVG_stacked_hosts.png",
       stacked, width = 18, height = 10, dpi = 300)

ggsave("Images/DVG Segment Plots - combined/DVG_stacked_hosts.svg",
       stacked, width = 18, height = 10, dpi = 300)

```

DVGS PER SEGMENT, FRACTIONAL SUPPORT PER SEGMENT (INSTEAD OF TOTAL)

```{r}
###############################################################
## Compute fractional support *within each segment per sample*
###############################################################

dvg_frac <- dvg_filt %>%
  group_by(SampleID, Segment) %>%
  mutate(
    segment_total_support = sum(mean_read_counts, na.rm = TRUE),
    frac_support_segment  = mean_read_counts / segment_total_support
  ) %>%
  ungroup()

###############################################
## 1. SEPARATE FRACTIONAL SUPPORT PLOTS
###############################################

for (segment in segs) {

  df1 <- dvg_frac %>% filter(Segment == segment)
  if (nrow(df1) == 0) next

  seg_start <- segment_ends[[segment]][1]
  seg_end   <- segment_ends[[segment]][2]

  delheight <- 0.01
  delwidth  <- 0.7

  # Dummy rows so all Hosts appear in legend/colors
  dummy_data <- expand.grid(
    SampleID="dummy",
    Host=names(colors),
    Segment=segment,
    BP=seg_start,
    RI=seg_start+1,
    frac_support_segment=0
  )

  df1_plot <- bind_rows(df1, dummy_data)

  p <- ggplot(df1_plot, aes(color = Host)) +

    geom_segment(aes(x=seg_start, xend=BP,
                     y=frac_support_segment, yend=frac_support_segment),
                 linewidth=0.2, linetype="dotted") +

    geom_segment(aes(x=BP, xend=BP,
                     y=frac_support_segment - delheight,
                     yend=frac_support_segment + delheight),
                 linewidth=delwidth) +

    geom_segment(aes(x=RI, xend=RI,
                     y=frac_support_segment - delheight,
                     yend=frac_support_segment + delheight),
                 linewidth=delwidth) +

    geom_segment(aes(x=RI, xend=seg_end,
                     y=frac_support_segment, yend=frac_support_segment),
                 linewidth=0.2, linetype="dotted") +

    annotate("text",
             x = seg_start + (seg_end - seg_start)/2,
             y = 0.48,
             label = segment,
             size = 7, fontface = "bold") +

    scale_color_manual(values = colors) +

    scale_x_continuous(limits = c(seg_start, seg_end),
                       breaks = seq(0, seg_end, 250),
                       expand = c(0,0)) +

    scale_y_continuous(limits=c(0,1),
                       breaks=seq(0,1,0.1),
                       expand=c(0,0)) +

    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      axis.text  = element_text(size=16),
      axis.title = element_text(size=20, face="bold"),
      legend.position="none",
      axis.line = element_line(color="black"),
      axis.ticks = element_line(color="black")
    ) +
    labs(
      x = "Genomic position",
      y = "Fraction of DVG reads"
    )

  ggsave(paste0("Images/DVG Fraction per Segment/DVGlines_frac_", segment, ".png"),
         p, width=8, height=5, dpi=300)
  ggsave(paste0("Images/DVG Fraction per Segment/DVGlines_frac_", segment, ".svg"),
         p, width=8, height=5, dpi=300)
}

###############################################
## 2. COMBINED FRACTIONAL SUPPORT PLOT
###############################################

library(patchwork)

segment_order <- segs

# Build combined DF
full_plot_df <- bind_rows(lapply(segment_order, function(seg) {
  df <- dvg_frac %>% filter(Segment == seg)
  if (nrow(df)==0) return(NULL)
  df$seg_start <- segment_ends[[seg]][1]
  df$seg_end   <- segment_ends[[seg]][2]
  df
}))

full_plot_df$Segment <- factor(full_plot_df$Segment,
                               levels = segment_order)

plot_seg_frac <- function(df_seg, seg_name, first=FALSE){

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]

  th <- 0.01   # tick half-height
  tw <- 0.7    # tick width

  ggplot(df_seg, aes(color=Host)) +
    geom_segment(aes(x=seg_start, xend=BP,
                     y=frac_support_segment, yend=frac_support_segment),
                 linewidth=0.2, linetype="dotted") +

    geom_segment(aes(x=BP, xend=BP,
                     y=frac_support_segment - th,
                     yend=frac_support_segment + th),
                 linewidth=tw) +

    geom_segment(aes(x=RI, xend=RI,
                     y=frac_support_segment - th,
                     yend=frac_support_segment + th),
                 linewidth=tw) +

    geom_segment(aes(x=RI, xend=seg_end,
                     y=frac_support_segment, yend=frac_support_segment),
                 linewidth=0.2, linetype="dotted") +

    scale_color_manual(values = colors) +

    scale_x_continuous(
      limits  = c(seg_start, seg_end),
      breaks  = seq(0, seg_end, 500),
      expand  = c(0,0)
    ) +

    scale_y_continuous(
      limits  = c(0, 1),
      breaks  = seq(0, 1, 0.1),
      expand  = c(0,0)
    ) +

    labs(
      x = seg_name,
      y = if (first) "Fraction of DVG reads" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      legend.position = "none",

      axis.text.x  = element_text(size=12, color="black"),
      axis.title.x = element_text(size=14, face="bold"),

      # y-axis only visible on first panel
      axis.line.y  = if(first) element_line(color="black", linewidth=1) else element_blank(),
      axis.ticks.y = if(first) element_line(color="black") else element_blank(),
      axis.text.y  = if(first) element_text(size=12, color="black") else element_blank(),
      axis.title.y = if(first) element_text(size=14, face="bold") else element_blank(),

      # x-axis always ON
      axis.line.x  = element_line(color="black", linewidth=1),
      axis.ticks.x = element_line(color="black"),

      axis.ticks.length = unit(4, "pt")
    )
}

plots <- lapply(seq_along(segment_order), function(i){
  seg <- segment_order[i]
  df_seg <- full_plot_df %>% filter(Segment==seg)
  plot_seg_frac(df_seg, seg, first=(i==1))
})

segment_lengths <- sapply(segment_ends, function(x) x[2] - x[1])
relative_widths <- segment_lengths[segment_order] / sum(segment_lengths)

combined_frac <- wrap_plots(plots, nrow=1, widths=relative_widths)

ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_fraction.png",
       combined_frac, width=18, height=6, dpi=300)
ggsave("Images/DVG Segment Plots - combined/DVG_all_segments_fraction.svg",
       combined_frac, width=18, height=6, dpi=300)

```
___________________________
___________________________
___GOOD, THIS IS THE ONE___
___________________________
___________________________
DVG HILLS SEPARATED BY HOST, Y axis is each Species, counted and stacked.

```{r}
############################################################
## 1. SETTINGS
############################################################

segment_order <- c("PB2","PB1","PA","HA","NP","NR","M","NS")

segment_ends <- list(
  PB2 = c(1, 2341),
  PB1 = c(1, 2341),
  PA  = c(1, 2233),
  HA  = c(1, 1778),
  NP  = c(1, 1565),
  NR  = c(1, 1433),
  M   = c(1, 1027),
  NS  = c(1,  890)
)

host_order <- c("BALBF","BALBM","BL6F","BL6M")

colors <- c(
  "BALBF" = "#F94040",
  "BALBM" = "#0000FF",
  "BL6F"  = "#C5944E",
  "BL6M"  = "#0F99B2"
)

y_max_stack <- 420   # fixed y limit

############################################################
## 2. PREP DATA FOR STACKED PLOT
############################################################
# dvg_filt is assumed to exist in your environment

dvg_stack <- dvg_filt %>%
  filter(!is.na(Segment),
         Segment %in% segment_order,
         !is.na(BP),
         !is.na(RI)) %>%
  mutate(
    Segment = factor(Segment, levels = segment_order),
    Host    = factor(Host,    levels = host_order),
    del_len = abs(RI - BP)
  ) %>%
  # one row per unique DVG within Segment/Host
  group_by(Segment, Host, cID_DI) %>%
  summarise(
    BP      = first(BP),
    RI      = first(RI),
    del_len = first(del_len),
    .groups = "drop"
  ) %>%
  group_by(Segment, Host) %>%
  arrange(desc(del_len), .by_group = TRUE) %>%
  mutate(stack_y = row_number()) %>%
  ungroup()

############################################################
## 3. FUNCTION: ONE SEGMENT, 4 HOST PANELS SIDE BY SIDE
############################################################

plot_segment_hosts <- function(seg_name, show_y = FALSE) {

  df_seg <- dvg_stack %>% filter(Segment == seg_name)
  if (nrow(df_seg) == 0) return(ggplot() + theme_void())

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]

  ggplot(df_seg, aes(color = Host)) +
    # start → BP (genome, hidden)
    geom_segment(aes(x = seg_start, xend = BP,
                     y = stack_y,    yend = stack_y),
                 linewidth = 0.1, alpha = 0.3, linetype = "solid") +
    # BP tick
    geom_segment(aes(x = BP, xend = BP,
                     y = stack_y - 1, yend = stack_y + 1),
                 linewidth = 0.3, alpha = 1) +
    # BP → RI (deletion)
    geom_segment(aes(x = BP, xend = RI,
                     y = stack_y, yend = stack_y),
                 linewidth = 0.1, alpha = 1, linetype = "solid") +
    # RI tick
    geom_segment(aes(x = RI, xend = RI,
                     y = stack_y - 1, yend = stack_y + 1),
                 linewidth = 0.3, alpha = 1) +
    # RI → end (genome, hidden)
    geom_segment(aes(x = RI, xend = seg_end,
                     y = stack_y, yend = stack_y),
                 linewidth = 0.1, alpha = 0.3, linetype = "solid") +

    facet_wrap(~ Host, nrow = 1, strip.position = "bottom") +
    theme(
      strip.text       = element_blank(),
      strip.background = element_blank()
    ) +

    scale_color_manual(values = colors, drop = FALSE) +
    (
      {
        seg_len <- seg_end - seg_start

        if (seg_len < 1600) {
          # short segments → one tick at 500
          scale_x_continuous(
            limits = c(seg_start, seg_end),
            breaks = 500,
            minor_breaks = NULL,
            expand = c(0, 0)
          )
        } else {
          # long segments → usual ticks
          scale_x_continuous(
            limits = c(seg_start, seg_end),
            breaks = seq(500, seg_end, by = 1000),
            minor_breaks = seq(0, seg_end, by = 250),
            expand = c(0, 0)
          )
        }
      }
    ) +

    scale_y_continuous(
      limits = c(0, y_max_stack),
      breaks = seq(0, y_max_stack, by = 100),
      expand = c(0, 0)
    ) +

    labs(
      x = seg_name,
      y = if (show_y) "Unique DVGs" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid      = element_blank(),
      legend.position = "none",
      panel.spacing.x = unit(0.1, "lines"),   # hosts close together

      axis.text.x  = element_text(size = 7, color = "black"),
      axis.title.x = element_text(size = 14, face = "bold"),
      axis.line.x  = element_line(color = "black"),
      axis.ticks.x = element_line(size = 0.4, color = "black"),
      axis.ticks.length.x = unit(3, "pt"),

      axis.text.y  = if (show_y) element_text(size = 10, color = "black") else element_blank(),
      axis.title.y = if (show_y) element_text(size = 14, face = "bold") else element_blank(),
      axis.ticks.y = if (show_y) element_line(color = "black") else element_blank(),
      axis.line.y  = if (show_y) element_line(color = "black") else element_blank(),

      plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
    ) +
    theme(
      strip.background.x = element_blank(),
      strip.text.x       = element_blank(),
      strip.background.y = element_blank(),
      strip.text.y       = element_blank()
    )
}

############################################################
## 4. BUILD ALL SEGMENT PANELS AND COMBINE
############################################################

segment_plots <- lapply(seq_along(segment_order), function(i) {
  seg <- segment_order[i]
  show_y <- (i == 1)   # only PB2 shows y axis
  plot_segment_hosts(seg, show_y = show_y)
})

# widen small segments a bit so there is visible space between segments
seg_lengths <- sapply(segment_ends, function(x) x[2] - x[1])
seg_lengths_padded <- seg_lengths[segment_order] + 1000
relative_widths <- seg_lengths_padded / sum(seg_lengths_padded)

final_plot <- wrap_plots(
  segment_plots,
  nrow   = 1,
  widths = relative_widths
)

############################################################
## 5. SAVE
############################################################

if (!dir.exists("Images/DVG Segment Plots - stacked"))
  dir.create("Images/DVG Segment Plots - stacked", recursive = TRUE)

ggsave("Images/DVG Segment Plots - stacked/DVG_stacked_species_byHost.png",
       final_plot, width = 16, height = 8, dpi = 300)

ggsave("Images/DVG Segment Plots - stacked/DVG_stacked_species_byHost.svg",
       final_plot, width = 16, height = 8, dpi = 300)

```

COLORED BACKGROUND

```{r}
############################################################
## 1. SETTINGS
############################################################

segment_order <- c("PB2","PB1","PA","HA","NP","NR","M","NS")

segment_ends <- list(
  PB2 = c(1, 2341),
  PB1 = c(1, 2341),
  PA  = c(1, 2233),
  HA  = c(1, 1778),
  NP  = c(1, 1565),
  NR  = c(1, 1433),
  M   = c(1, 1027),
  NS  = c(1,  890)
)

host_order <- c("BALBF","BALBM","BL6F","BL6M")

colors <- c(
  "BALBF" = "#F94040",
  "BALBM" = "#0000FF",
  "BL6F"  = "#C5944E",
  "BL6M"  = "#0F99B2"
)

y_max_stack <- 420   # fixed y limit

############################################################
## 2. PREP DATA FOR STACKED PLOT
############################################################
# dvg_filt is assumed to exist in your environment

dvg_stack <- dvg_filt %>%
  filter(!is.na(Segment),
         Segment %in% segment_order,
         !is.na(BP),
         !is.na(RI)) %>%
  mutate(
    Segment = factor(Segment, levels = segment_order),
    Host    = factor(Host,    levels = host_order),
    del_len = abs(RI - BP)
  ) %>%
  # one row per unique DVG within Segment/Host
  group_by(Segment, Host, cID_DI) %>%
  summarise(
    BP      = first(BP),
    RI      = first(RI),
    del_len = first(del_len),
    .groups = "drop"
  ) %>%
  group_by(Segment, Host) %>%
  arrange(desc(del_len), .by_group = TRUE) %>%
  mutate(stack_y = row_number()) %>%
  ungroup()

############################################################
## 3. FUNCTION: ONE SEGMENT, 4 HOST PANELS SIDE BY SIDE
############################################################

plot_segment_hosts <- function(seg_name, show_y = FALSE) {

  df_seg <- dvg_stack %>% filter(Segment == seg_name)
  if (nrow(df_seg) == 0) return(ggplot() + theme_void())

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]

  ggplot(df_seg, aes(color = Host)) +
    # start → BP (genome, hidden)
    geom_segment(aes(x = seg_start, xend = BP,
                     y = stack_y,    yend = stack_y),
                 linewidth = 0.2, alpha = 0.8, linetype = "solid") +
    # BP tick
    geom_segment(aes(x = BP, xend = BP,
                     y = stack_y - 1, yend = stack_y + 1),
                 linewidth = 0.3, alpha = 1) +
    # BP → RI (deletion)
    geom_segment(aes(x = BP, xend = RI,
                     y = stack_y, yend = stack_y),
                 linewidth = 0.2, alpha = 0.2, linetype = "solid") +
    # RI tick
    geom_segment(aes(x = RI, xend = RI,
                     y = stack_y - 1, yend = stack_y + 1),
                 linewidth = 0.3, alpha = 1) +
    # RI → end (genome, hidden)
    geom_segment(aes(x = RI, xend = seg_end,
                     y = stack_y, yend = stack_y),
                 linewidth = 0.2, alpha = 0.8, linetype = "solid") +

    facet_wrap(~ Host, nrow = 1, strip.position = "bottom") +
    theme(
      strip.text       = element_blank(),
      strip.background = element_blank()
    ) +

    scale_color_manual(values = colors, drop = FALSE) +
    (
      {
        seg_len <- seg_end - seg_start

        if (seg_len < 1600) {
          # short segments → one tick at 500
          scale_x_continuous(
            limits = c(seg_start, seg_end),
            breaks = 500,
            minor_breaks = NULL,
            expand = c(0, 0)
          )
        } else {
          # long segments → usual ticks
          scale_x_continuous(
            limits = c(seg_start, seg_end),
            breaks = seq(500, seg_end, by = 1000),
            minor_breaks = seq(0, seg_end, by = 250),
            expand = c(0, 0)
          )
        }
      }
    ) +

    scale_y_continuous(
      limits = c(0, y_max_stack),
      breaks = seq(0, y_max_stack, by = 100),
      expand = c(0, 0)
    ) +

    labs(
      x = seg_name,
      y = if (show_y) "Unique DVGs" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid      = element_blank(),
      legend.position = "none",
      panel.spacing.x = unit(0.1, "lines"),   # hosts close together

      axis.text.x  = element_text(size = 7, color = "black"),
      axis.title.x = element_text(size = 14, face = "bold"),
      axis.line.x  = element_line(color = "black"),
      axis.ticks.x = element_line(size = 0.4, color = "black"),
      axis.ticks.length.x = unit(3, "pt"),

      axis.text.y  = if (show_y) element_text(size = 10, color = "black") else element_blank(),
      axis.title.y = if (show_y) element_text(size = 14, face = "bold") else element_blank(),
      axis.ticks.y = if (show_y) element_line(color = "black") else element_blank(),
      axis.line.y  = if (show_y) element_line(color = "black") else element_blank(),

      plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
    ) +
    theme(
      strip.background.x = element_blank(),
      strip.text.x       = element_blank(),
      strip.background.y = element_blank(),
      strip.text.y       = element_blank()
    )
}

############################################################
## 4. BUILD ALL SEGMENT PANELS AND COMBINE
############################################################

segment_plots <- lapply(seq_along(segment_order), function(i) {
  seg <- segment_order[i]
  show_y <- (i == 1)   # only PB2 shows y axis
  plot_segment_hosts(seg, show_y = show_y)
})

# widen small segments a bit so there is visible space between segments
seg_lengths <- sapply(segment_ends, function(x) x[2] - x[1])
seg_lengths_padded <- seg_lengths[segment_order] + 1000
relative_widths <- seg_lengths_padded / sum(seg_lengths_padded)

final_plot <- wrap_plots(
  segment_plots,
  nrow   = 1,
  widths = relative_widths
)

############################################################
## 5. SAVE
############################################################

if (!dir.exists("Images/DVG Segment Plots - stacked"))
  dir.create("Images/DVG Segment Plots - stacked", recursive = TRUE)

ggsave("Images/DVG Segment Plots - stacked/DVG_stacked_species_reversed.png",
       final_plot, width = 16, height = 8, dpi = 300)

ggsave("Images/DVG Segment Plots - stacked/DVG_stacked_species_reversed.svg",
       final_plot, width = 16, height = 8, dpi = 300)
```


ONLY ALTERNATE DELETION TYPES TO SEE WHAT THEY LOOK LIKE

```{r}
############################################################
## 1. SETTINGS
############################################################

segment_order <- c("PB2","PB1","PA","HA","NP","NR","M","NS")

segment_ends <- list(
  PB2 = c(1, 2341),
  PB1 = c(1, 2341),
  PA  = c(1, 2233),
  HA  = c(1, 1778),
  NP  = c(1, 1565),
  NR  = c(1, 1433),
  M   = c(1, 1027),
  NS  = c(1,  890)
)

host_order <- c("BALBF","BALBM","BL6F","BL6M")

colors <- c(
  "BALBF" = "#F94040",
  "BALBM" = "#0000FF",
  "BL6F"  = "#C5944E",
  "BL6M"  = "#0F99B2"
)

y_max_stack <- 420   # fixed y limit

############################################################
## 2. PREP DATA FOR STACKED PLOT
############################################################
# dvg_filt is assumed to exist in your environment

dvg_stack <- dvg_filt %>%
  filter(!is.na(Segment),
         Segment %in% segment_order,
         !is.na(BP),
         !is.na(RI),
         DVG_type != "Deletion_forward") %>%   # <-- FILTER OUT 3cb/sb HERE
  mutate(
    Segment = factor(Segment, levels = segment_order),
    Host    = factor(Host,    levels = host_order),
    del_len = abs(RI - BP)
  ) %>%
  # one row per unique DVG within Segment/Host
  group_by(Segment, Host, cID_DI) %>%
  summarise(
    BP      = first(BP),
    RI      = first(RI),
    del_len = first(del_len),
    .groups = "drop"
  ) %>%
  group_by(Segment, Host) %>%
  arrange(desc(del_len), .by_group = TRUE) %>%
  mutate(stack_y = row_number()) %>%
  ungroup()

############################################################
## 3. FUNCTION: ONE SEGMENT, 4 HOST PANELS SIDE BY SIDE
############################################################

plot_segment_hosts <- function(seg_name, show_y = FALSE) {

  df_seg <- dvg_stack %>% filter(Segment == seg_name)
  if (nrow(df_seg) == 0) return(ggplot() + theme_void())

  seg_start <- segment_ends[[seg_name]][1]
  seg_end   <- segment_ends[[seg_name]][2]

  ggplot(df_seg, aes(color = Host)) +
    # start → BP (genome, hidden)
    geom_segment(aes(x = seg_start, xend = BP,
                     y = stack_y,    yend = stack_y),
                 linewidth = 0.1, alpha = 0.3, linetype = "solid") +
    # BP tick
    geom_segment(aes(x = BP, xend = BP,
                     y = stack_y - 1, yend = stack_y + 1),
                 linewidth = 0.3, alpha = 1) +
    # BP → RI (deletion)
    geom_segment(aes(x = BP, xend = RI,
                     y = stack_y, yend = stack_y),
                 linewidth = 0.1, alpha = 1, linetype = "solid") +
    # RI tick
    geom_segment(aes(x = RI, xend = RI,
                     y = stack_y - 1, yend = stack_y + 1),
                 linewidth = 0.3, alpha = 1) +
    # RI → end (genome, hidden)
    geom_segment(aes(x = RI, xend = seg_end,
                     y = stack_y, yend = stack_y),
                 linewidth = 0.1, alpha = 0.3, linetype = "solid") +

    facet_wrap(~ Host, nrow = 1, strip.position = "bottom") +
    theme(
      strip.text       = element_blank(),
      strip.background = element_blank()
    ) +

    scale_color_manual(values = colors, drop = FALSE) +
    (
      {
        seg_len <- seg_end - seg_start

        if (seg_len < 1600) {
          # short segments → one tick at 500
          scale_x_continuous(
            limits = c(seg_start, seg_end),
            breaks = 500,
            minor_breaks = NULL,
            expand = c(0, 0)
          )
        } else {
          # long segments → usual ticks
          scale_x_continuous(
            limits = c(seg_start, seg_end),
            breaks = seq(500, seg_end, by = 1000),
            minor_breaks = seq(0, seg_end, by = 250),
            expand = c(0, 0)
          )
        }
      }
    ) +

    scale_y_continuous(
      limits = c(0, y_max_stack),
      breaks = seq(0, y_max_stack, by = 100),
      expand = c(0, 0)
    ) +

    labs(
      x = seg_name,
      y = if (show_y) "Unique DVGs" else NULL
    ) +

    theme_minimal() +
    theme(
      panel.grid      = element_blank(),
      legend.position = "none",
      panel.spacing.x = unit(0.1, "lines"),   # hosts close together

      axis.text.x  = element_text(size = 7, color = "black"),
      axis.title.x = element_text(size = 14, face = "bold"),
      axis.line.x  = element_line(color = "black"),
      axis.ticks.x = element_line(size = 0.4, color = "black"),
      axis.ticks.length.x = unit(3, "pt"),

      axis.text.y  = if (show_y) element_text(size = 10, color = "black") else element_blank(),
      axis.title.y = if (show_y) element_text(size = 14, face = "bold") else element_blank(),
      axis.ticks.y = if (show_y) element_line(color = "black") else element_blank(),
      axis.line.y  = if (show_y) element_line(color = "black") else element_blank(),

      plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
    ) +
    theme(
      strip.background.x = element_blank(),
      strip.text.x       = element_blank(),
      strip.background.y = element_blank(),
      strip.text.y       = element_blank()
    )
}

############################################################
## 4. BUILD ALL SEGMENT PANELS AND COMBINE
############################################################

segment_plots <- lapply(seq_along(segment_order), function(i) {
  seg <- segment_order[i]
  show_y <- (i == 1)   # only PB2 shows y axis
  plot_segment_hosts(seg, show_y = show_y)
})

# widen small segments a bit so there is visible space between segments
seg_lengths <- sapply(segment_ends, function(x) x[2] - x[1])
seg_lengths_padded <- seg_lengths[segment_order] + 1000
relative_widths <- seg_lengths_padded / sum(seg_lengths_padded)

final_plot <- wrap_plots(
  segment_plots,
  nrow   = 1,
  widths = relative_widths
)

############################################################
## 5. SAVE
############################################################

if (!dir.exists("Images/DVG Segment Plots - stacked"))
  dir.create("Images/DVG Segment Plots - stacked", recursive = TRUE)

ggsave("Images/DVG Segment Plots - stacked/DVG_stacked_species_rare.png",
       final_plot, width = 16, height = 8, dpi = 300)

ggsave("Images/DVG Segment Plots - stacked/DVG_stacked_species_rare.svg",
       final_plot, width = 16, height = 8, dpi = 300)

```




_____________________________
_____________________________
___SEPARATED TYPES OF DVGs___
_____________________________


DVG PER TYPE GRAPH WITH BARS, SAMPLES WITH NO DVGs NOT INCLUDED.
```{r}

# Compute mean per group (bar heights)
bar_means <- dvg_type_counts %>%
  group_by(DVG_type, Host) %>%
  summarise(mean_n = mean(n), .groups="drop")

# Order DVG types logically
dvg_order <- c("Deletion_forward", "Deletion_reverse", "3cb/sb", "Insertion_forward")

dvg_type_counts$DVG_type <- factor(dvg_type_counts$DVG_type, levels=dvg_order)
bar_means$DVG_type       <- factor(bar_means$DVG_type,       levels=dvg_order)
# Replace zeros for log plotting
dvg_plot <- dvg_type_counts %>%
  mutate(n_plot = ifelse(n == 0, 0.3, n))

host_colors <- c(
  "BALBF"="#F94040",
  "BALBM"="#0000FF",
  "BL6F" ="#C5944E",
  "BL6M" ="#0F99B2"
)

dodge <- position_dodge(width = 0.7)

p <- ggplot(dvg_plot, aes(x = DVG_type, y = n_plot, fill = Host)) +

  # Mean bars (solid bars, fully visible)
  geom_bar(
    stat = "summary",
    fun = "mean",
    position = dodge,
    width = 0.6,
    alpha = 0.15          # transparent bars
  ) +

  # SE whiskers
  geom_errorbar(
    stat = "summary",
    fun.data = mean_se,
    position = dodge,
    width = 0.4,
    color = "black",
    linewidth = 0.5
  ) +

  # Individual points
  geom_jitter(
    aes(color = Host),
    position = position_jitterdodge(
      jitter.width = 0.2,
      dodge.width = 0.7
    ),
    size = 2.2,
    alpha = 0.9
  ) +

  scale_fill_manual(values = host_colors) +
  scale_color_manual(values = host_colors) +

  # Log scale but show zero
  scale_y_log10(
    breaks = c(0.3, 1, 3, 10, 30, 100, 300),
    labels = c("0", "1", "3", "10", "30", "100", "300")
  ) +

  labs(x = "", y = "Unique DVGs (log scale)") +

  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color="black"),
    legend.position = "none"
  )

p


ggsave("Images/DVG_Types/DVGtype_grouped_meanSE_Bars.png",
       p, width = 13, height = 6, dpi = 300)
```


NO BARS, SAMPLES WITH ZERO INCLUDED. UNFORTUNATELY, CAN'T PLOT MEANS FOR SOME STUPID REASON.

```{r}
## ============================================================
## DVG TYPE COUNTS FOR PLOT
## ============================================================

# Types to keep (omit Insertions)
plot_types <- c("Deletion_forward", "Deletion_reverse", "3cb/sb")

# Custom labels for the x axis
custom_labels <- c(
  "Deletion_forward"  = "Forward deletions",
  "Deletion_reverse"  = "Reverse deletions",
  "3cb/sb"            = "3' copy-back/snap-back"
)

# All samples that passed coverage filtering and remain in `dvg`
# (this is your "good samples" universe)
all_samples <- dvg %>%
  distinct(SampleID, Host, HostStrain, Sex) %>%
  filter(!is.na(Host), !is.na(HostStrain), !is.na(Sex))

# Restrict dvg_type_counts to the types we actually want to plot
dvg_type_counts_plot <- dvg_type_counts %>%
  filter(DVG_type %in% plot_types)

# Expand to all SampleID × DVG_type combinations in the good-sample universe,
# then left-join counts and fill missing with zero.
dvg_plot <- all_samples %>%
  tidyr::crossing(DVG_type = plot_types) %>%
  left_join(dvg_type_counts_plot,
            by = c("SampleID", "Host", "HostStrain", "Sex", "DVG_type")) %>%
  mutate(
    n = replace_na(n, 0),
    # factor with desired order and labels
    DVG_type = factor(DVG_type,
                      levels = plot_types,
                      labels = custom_labels),
    # value used for plotting on log scale
    n_plot = ifelse(n == 0, 0.3, n)
  )

bar_rect <- dvg_plot %>%
  group_by(DVG_type, Host) %>%
  summarise(
    ymin = 0.3,             # <-- your chosen zero
    ymax = mean(n_plot),    # <-- bar height
    .groups = "drop"
  ) %>%
  mutate(
    x = as.numeric(DVG_type),
    xmin = x + c(-0.3, -0.1, 0.1, 0.3)[as.numeric(Host)],  # 4 hosts spaced
    xmax = x + c(-0.1, 0.1, 0.3, 0.5)[as.numeric(Host)]
  )


## ============================================================
## PLOT
## ============================================================

host_colors <- c(
  "BALBF" = "#F94040",
  "BALBM" = "#0000FF",
  "BL6F"  = "#C5944E",
  "BL6M"  = "#0F99B2"
)

dodge <- position_dodge(width = 0.7)

p <- ggplot(dvg_plot, aes(x = DVG_type, y = n_plot, fill = Host)) +

  # Mean bars (semi transparent)
geom_rect(
  data = bar_rect,
  aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  fill = "black",
  alpha = 0.80,
  inherit.aes = FALSE
) +

  # SE whiskers
  geom_errorbar(
    stat      = "summary",
    fun.data  = mean_se,
    position  = dodge,
    width     = 0.4,
    color     = "black",
    linewidth = 0.5
  ) +

  # Individual sample points
  geom_jitter(
    aes(color = Host),
    position = position_jitterdodge(
      jitter.width = 0.2,
      dodge.width  = 0.7
    ),
    size  = 2.2,
    alpha = 0.9
  ) +

  scale_fill_manual(values = host_colors) +
  scale_color_manual(values = host_colors) +

  # Log scale with 0.3 shown as "0"
  scale_y_log10(
    breaks = c(0.3, 1, 3, 10, 30, 100, 300),
    labels = c("0",  "1", "3", "10", "30", "100", "300"),
      ) +

  labs(x = "", y = "Unique DVGs (log scale)") +

  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line        = element_line(color = "black"),
    axis.ticks       = element_line(color = "black"),
    legend.position  = "none"
  )

p

ggsave("Images/DVG_Types/DVGtype_grouped_meanSE_transparentBars.png",
       p, width = 10, height = 6, dpi = 300)

```



