kni---
title: "Survival tests for host differences"
author: "RC"
date: "2025-12-20"
output: html_document
---
Initially did only Cox Proportional hazards to account for multiple variables including strain. Since we are doing things a bit differently now, doing logrank as well. need to decide what to do.

I want to change every csv summary into a sink of the print of the test result, for cox and also for lmms. Because that reports the test that is being done on the data, making it much easier to understand what is being tested.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(survival)
library(tidyverse)
```

Define function for longer pvalue numbers in report
```{r}
pvalue <- function(x, ...) UseMethod("pvalue")
pvalue.survdiff <- function (x, ...) {
  if (length(x$n) == 1) {
    df <- 1
    pval <- pchisq(x$chisq, 1, lower.tail = FALSE)
  } else {
    if (is.matrix(x$obs)) {
      otmp <- rowSums(x$obs)
      etmp <- rowSums(x$exp)
    } else {
      otmp <- x$obs
      etmp <- x$exp
    }
    df <- sum(etmp > 0) - 1
    pval <- pchisq(x$chisq, df, lower.tail = FALSE)
  }
  list(chisq = x$chisq, p.value = pval, df = df)
  cat("Chi-squared:", x$chisq, "\nP-value:", pval, "\nDegrees of Freedom:", df, "\n")
}

```


Load data
```{r}
survdata<- read.csv("Groupedvirulence.master.csv", sep=',', header=T)
```

Make dataframes
```{r}
#All the data, including unfamiliar
survdata2 <- subset(survdata, Treatment!="1XP" & Treatment !="STOCK") 
#Only familiar infections
survdatafam <- subset(survdata, Treatment!="1XP" & Treatment!="STOCK" & Familiarity=="FAM")

#Set base levels
survdatafam$Host <- relevel(factor(survdatafam$Host), ref = "BALBF")
survdatafam$Treatment <- relevel(factor(survdatafam$Treatment), ref = "1X")

```

___HOST DIFFERENCES (Familiar)___
Logrank for comparisons between unpassaged and passaged. Change hosts as needed.
```{r}
# Subset your data
survdata3 <- subset(survdata, Treatment!="1XP" & Familiarity=="FAM")
host_levels <- c("BALBF", "BL6F", "BALBM")
output_file <- "logrank_results_by_host.txt"

sink(output_file)

for (host in host_levels) {

  df_host <- subset(survdata3, Host == host)

  surv_obj <- Surv(df_host$TTD, df_host$Event)

  logrank_result <- survdiff(surv_obj ~ factor(Status), data = df_host)

  cat("Log-rank test results for Host = ", host, ":\n", sep = "")
  print(logrank_result)

  # exact-ish p value from chisq
  p_val <- 1 - pchisq(logrank_result$chisq, df = length(logrank_result$n) - 1)
  cat("\nChi-squared: ", logrank_result$chisq,
      "\nDegrees of freedom: ", length(logrank_result$n) - 1,
      "\nP-value: ", format.pval(p_val, digits = 4, eps = 1e-16),
      "\n\n", sep = "")
}

sink()

```

Cox proportional hazards for comparisons between hosts:

```{r}
# Define the host levels and the output file name
host_levels <- c("BALBF", "BL6F", "BALBM")
output_file <- "CoxPH_Host_results_combined.txt"

# Open the output file for writing
sink(output_file)

# Iterate over the host levels
for (host in host_levels) {
  
  # Relevel Host variable to set the current host as the reference level
  survdatafam$Host <- relevel(factor(survdatafam$Host), ref = host)
  
  # Fit the CoxPH model for the current host level
  cox_model <- coxph(Surv(TTD, Event) ~ Host, data = survdatafam)
  
  # Print the result to the file
  cat("\n\nCox Proportional Hazards Model for Host = ", host, " (reference = ", host, "):\n", sep = "")
  print(cox_model)
}

# Close the file
sink()
```

With confidence intervals
```{r}
host_levels <- c("BALBF", "BL6F", "BALBM")
output_file <- "CoxPH_Host_results_with_CI.txt"

sink(output_file)

for (host in host_levels) {

  # Set reference level without permanently overwriting the column
  df <- survdatafam
  df$Host <- relevel(factor(df$Host), ref = host)

  cox_model <- coxph(Surv(TTD, Event) ~ Host, data = df)
  s <- summary(cox_model)

  cat("\n\nCox Proportional Hazards Model (reference = ", host, "):\n", sep = "")
  cat("n = ", s$n, ", events = ", s$nevent, "\n", sep = "")

  # Coefficients table (coef, HR, SE, z, p)
  cat("\nCoefficients:\n")
  print(s$coefficients)

  # HR and 95% CI (plus exp(coef) again)
  cat("\nHazard ratios and 95% CI:\n")
  # Columns: exp(coef), exp(-coef), lower .95, upper .95
  print(s$conf.int)

  # Optional: global tests
  cat("\nGlobal tests:\n")
  cat("Likelihood ratio test = ", s$logtest["test"], " on ",
      s$logtest["df"], " df, p = ", s$logtest["pvalue"], "\n", sep = "")
  cat("Wald test = ", s$waldtest["test"], " on ",
      s$waldtest["df"], " df, p = ", s$waldtest["pvalue"], "\n", sep = "")
  cat("Score (logrank) test = ", s$sctest["test"], " on ",
      s$sctest["df"], " df, p = ", s$sctest["pvalue"], "\n", sep = "")
}

sink()
```
