---
title: "Pooling Survival"
author: "RC"
date: "2026-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(survival)
library(tidyverse)
```

```{r}
pvalue <- function(x, ...) UseMethod("pvalue")
pvalue.survdiff <- function (x, ...) {
  if (length(x$n) == 1) {
    df <- 1
    pval <- pchisq(x$chisq, 1, lower.tail = FALSE)
  } else {
    if (is.matrix(x$obs)) {
      otmp <- rowSums(x$obs)
      etmp <- rowSums(x$exp)
    } else {
      otmp <- x$obs
      etmp <- x$exp
    }
    df <- sum(etmp > 0) - 1
    pval <- pchisq(x$chisq, df, lower.tail = FALSE)
  }
  list(chisq = x$chisq, p.value = pval, df = df)
  cat("Chi-squared:", x$chisq, "\nP-value:", pval, "\nDegrees of Freedom:", df, "\n")
}

```

Load data
```{r}
survdata<- read.csv("Groupedvirulence.master.csv", sep=',', header=T)
```

Make dataframes
```{r}
#All the data, including unfamiliar
survdata2 <- subset(survdata, Treatment!="1XP" & Treatment !="STOCK") 
#Only familiar infections
survdatafam <- subset(survdata, Treatment!="1XP" & Treatment!="STOCK" & Familiarity=="FAM")

#Set base levels
survdatafam$Host <- relevel(factor(survdatafam$Host), ref = "BALBF")
survdatafam$Treatment <- relevel(factor(survdatafam$Treatment), ref = "1X")

```

___POOLING DIFFERENCES___
Cox proportional hazards for comparisons between pooling treatments, including host as a fixed effect (but not looking at interactions):

```{r}
# Fit a CoxPH model for Treatment only
survdatafam$Host <- relevel(factor(survdatafam$Host), ref = "BALBF")
survdatafam$Treatment <- relevel(factor(survdatafam$Treatment), ref = "1X")
cox_model <- coxph(Surv(TTD, Event) ~ Treatment + Host, data = survdatafam)

# Print the summary of the model to see the results
summary(cox_model)

# Save to txt (no csv possible)
sink("CoxPH-Treatment.txt")
print(cox_model)
sink()

```




___1X vs 1XP___
```{r}
survdatafem <- subset(survdata, Treatment=="1X" | Treatment=="1XP") #Only 1X and 1XP
survdatafem <- subset(survdatafem, Sex=="F")
survdatafemfam <- subset(survdatafem, Familiarity == "FAM")
```

Does 1XP increase mortality in general?
```{r}
# Convert Test to a factor
survdatafem$Test <- factor(survdatafem$Test)

#Relevel 
survdatafem$Test <- relevel(survdatafem$Test, ref="BL6F.FAM")

#Run analyses
survdatafem$TTD
table(survdatafem$Event,survdatafem$TTD)
Surv(survdatafem$TTD,survdatafem$Event)



cox <- coxph(Surv(survdatafem$TTD,survdatafem$Event) ~ factor(Treatment), data=survdatafem)
summary(cox)

#Save
sink("coxph-1XP-female-only.txt")
print(cox)
sink()
```
Yes


Does 1XP increase mortality in familiar more than unfamiliar?
```{r}
# Convert Test to a factor
survdatafem$Test <- factor(survdatafem$Test)

#Relevel 
survdatafem$Test <- relevel(survdatafem$Test, ref="BL6F.FAM")

#Run analyses
survdatafem$TTD
table(survdatafem$Event,survdatafem$TTD)
Surv(survdatafem$TTD,survdatafem$Event)



cox <- coxph(Surv(survdatafem$TTD,survdatafem$Event) ~ factor(Treatment) * factor(Familiarity), data=survdatafem)
summary(cox)

sink("coxph-1XP-treatment-familiarity.txt")
print(cox)
sink()
```
No, but that's because the increase in mortality makes them equal. So it actually negates the tradeoff of the bl6 into balb.

Complete test: separated by host and strain.
```{r}
# Convert Test to a factor
survdatafem$Test <- factor(survdatafem$Test)
survdatafem$Treatment <- factor(survdatafem$Treatment)
#Relevel 
survdatafem$Test <- relevel(survdatafem$Test, ref="BL6F.FAM")
survdatafem$Treatment <- relevel(survdatafem$Treatment, ref="1X")

#Run analyses
survdatafem$TTD
table(survdatafem$Event,survdatafem$TTD)
Surv(survdatafem$TTD,survdatafem$Event)



cox <- coxph(Surv(survdatafem$TTD,survdatafem$Event) ~ factor(Test) * factor(Treatment), data=survdatafem)
summary(cox)

sink("coxph-1XP-test.txt")
print(cox)
sink()
```

This shows that 1XP causes treatments to be similarly high virulence. But not as good test as above.