---
title: "Model Test"
author: "RC"
date: "2024-02-28"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, include = FALSE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(MuMIn)
```

Load data
```{r}
alldata <- read.csv("groupedvirulence7days.csv", sep=',', header=T)
```


WHAT ARE THE QUESTIONS?

1. How does each virus do in familiar vs unfamiliar settings?
  BALBF Fam vs BALBF unfam vs BL6F Fam vs BL6F Unfam. Agnostic to pooling treatment but without 1XP because otherwise I'm biasing, as the same lineages infect more than once.

2. Is one virus overall more virulent than the other?
  BALBF-adapted vs BL6-adapted, agnostic to treatment and host of infection.

3. Is one host more susceptible overall than the other?

4. How does treatment affect virulence in the context of familiar vs unfamiliar?
If viral diversity matters when infecting a new host, is 1XP more virulent?
5.  Test 1XP vs 1X and also 4X, agnostic to familiarity but also in the context of host.

5. F-test for variance of balb infections vs bl6 infections? Maybe have to normalize for mean weight loss per day, then look at spread.

##### BALB_FAM vs BALB_UNFAM vs BL6FAM vs BL6UNFAM

```{r BALB_FAM vs BALB_UNFAM vs BL6FAM vs BL6UNFAM}
# Remove data that is not relevant
tradeoffs <- subset(alldata, Treatment != "STOCK" & Treatment != "1XP" & Sex == "F")

# Scale and center
tradeoffs$Day <- scale(tradeoffs$Day, center = TRUE, scale = TRUE)

# Create a new column with Host and familiarity
tradeoffs$HostFam <- paste0(tradeoffs$Host,"_",tradeoffs$Familiarity)
tradeoffs$HostFam <- as.factor(tradeoffs$HostFam)

# Levels to iterate over
host_fams <- c("BALBF_FAM", "BALBF_UNF", "BL6F_FAM", "BL6F_UNF")

# Open the file for writing all summaries
sink("LMM-Tradeoffs_summary.txt")

# Iterate over each level
for (level in host_fams) {
  
  # Relevel HostFam
  tradeoffs$HostFam <- relevel(factor(tradeoffs$HostFam), ref = level)
  
  # Run model
  model.tradeoffs <- lmer(Pct ~ Day * HostFam + (Day|Line/ID), data = tradeoffs, REML = TRUE)
  
  # Write the title indicating the level
  cat("\n\n##############################################\n")
  cat("            Leveled on", level, "\n")
  cat("##############################################\n\n")
  
  # Output summaries to the file
  print(summary(model.tradeoffs))
}

# Save and close the file
sink()

```

Ok, so only significant difference is BALBF_UNF vs BL6F_FAM /BL6F_UNF.
BALB FAM vs BL6 (fam or unfam) are marginally significant.

That means that BL6 getting the balb virus lose the most weight per day, BL6-adapted virus causes less weight loss in both balb and bl6F. So the host of adaptation is more determinant for virulence than the host of infection.


What if we combine the data to see overall virulence of BALBF vs BL6F across familiar and unfamiliar hosts?

####BALB-adapted vs BL6-adapted

```{r BALB-adapted vs BL6-adapted}
#Remove data that is not relevant
overallvir <- subset(alldata, Treatment !="STOCK" & Sex == "F" & Treatment !="1XP" )

#Level of 1X and familiar

overallvir$Host <- relevel(factor(overallvir$Host), ref = "BALBF")
overallvir$Familiarity <- relevel(factor(overallvir$Familiarity), ref = "FAM")

#Scale and center.
overallvir$Day <- scale(overallvir$Day, center = TRUE, scale = TRUE)

#Run model
model.overallvir <- lmer(Pct ~ Day * Host + (Day|Line/ID), data = overallvir, REML = TRUE)

summary(model.overallvir)

# Open the file for writing
sink("LMM-Overall_virulence_BALBF_vs_BL6F.txt")

# Output summaries to the file
summary(model.overallvir)
# Save
sink()

```

But this includes 1X and 1XP, which are the same lineages, so perhaps not a fair comparison. Maybe let's do just 1X and 4X.

##### 1) Tradeoffs NO 1XP, NO 2X

```{rTradeoffs NO 1XP, NO 2X}
# Remove data that is not relevant
tradeoffs <- subset(alldata, Treatment != "STOCK" & Treatment != "2X" & Treatment != "1XP" & Sex == "F")

# Scale and center
tradeoffs$Day <- scale(tradeoffs$Day, center = TRUE, scale = TRUE)

# Create a new column with Status and Host -> Type
tradeoffs$HostFam <- paste0(tradeoffs$Host,"_",tradeoffs$Familiarity)
tradeoffs$HostFam <- as.factor(tradeoffs$HostFam)

# Levels to iterate over
host_fams <- c("BALBF_FAM", "BALBF_UNF", "BL6F_FAM", "BL6F_UNF")

# Open the file for writing all summaries
sink("LMM-Tradeoffs_1X&4X_summary.txt")

# Iterate over each level
for (level in host_fams) {
  
  # Relevel HostFam
  tradeoffs$HostFam <- relevel(factor(tradeoffs$HostFam), ref = level)
  
  # Run model
  model.tradeoffs <- lmer(Pct ~ Day * HostFam + (Day|Line/ID), data = tradeoffs, REML = TRUE)
  
  # Write the title indicating the level
  cat("\n\n##############################################\n")
  cat("            Leveled on", level, "\n")
  cat("##############################################\n\n")
  
  # Output summaries to the file
  print(summary(model.tradeoffs))
}

# Save and close the file
sink()

```



#### 2) BALB-adapted vs BL6-adapted NO 1XP. Significant. Whew.

```{rBALB-adapted vs BL6-adapted NO 1XP}
#Remove data that is not relevant
overallvir <- subset(alldata, Treatment !="STOCK" & Treatment != "2X" & Treatment != "1XP" & Sex == "F")

#Level of 1X and familiar

overallvir$Host <- relevel(factor(overallvir$Host), ref = "BALBF")
overallvir$Familiarity <- relevel(factor(overallvir$Familiarity), ref = "FAM")

#Scale and center.
overallvir$Day <- scale(overallvir$Day, center = TRUE, scale = TRUE)

#Run model
model.overallvir <- lmer(Pct ~ Day * Host + (Day|Line/ID), data = overallvir, REML = TRUE)

summary(model.overallvir)

# Open the file for writing
sink("LMM-1X&4X_virulence_BALBF_vs_BL6F.txt")

# Output summaries to the file
summary(model.overallvir)
# Save
sink()

```

###### 3) Is one host more susceptible overall than the other? HostStrain BALB/c vs C57BL/6 (combined viruses in each host) NO!

```{r HostStrain BALB/c vs C57BL/6 DELETE}
#Remove data that is not relevant
overallvir <- subset(alldata, Treatment !="STOCK" & Treatment != "2X" & Treatment != "1XP" & Sex == "F")

#Scale and center.
overallvir$Day <- scale(overallvir$Day, center = TRUE, scale = TRUE)

#Run model
model.overallsusc <- lmer(Pct ~ Day * HostStrain + (Day|Line/ID), data = overallvir, REML = TRUE)

summary(model.overallsusc)

# Open the file for writing
sink("LMM-1X&4X_susceptibility_BALBC_vs_BL6.txt")

# Output summaries to the file
summary(model.overallsusc)
# Save
sink()

```


Perhaps a repetitive question but is one host strain more susceptible to the familiar virus? What about the unfamiliar virus?

```{r STRAIN MORE SUSCEPTIBLE TO FAM - DELETE}
#Remove data that is not relevant
overallvir <- subset(alldata, Treatment !="STOCK" & Treatment != "2X" & Treatment != "1XP" & Sex == "F")

#Scale and center.
overallvir$Day <- scale(overallvir$Day, center = TRUE, scale = TRUE)

#overallvir <- subset(overallvir, Familiarity == "FAM")
overallvir <- subset(overallvir, Familiarity == "UNF")


#Run model
model.overallsusc <- lmer(Pct ~ Day * HostStrain + (Day|Line/ID), data = overallvir, REML = TRUE)

summary(model.overallsusc)

# Open the file for writing
sink("LMM-1X&4X_susceptibility_BALBC_vs_BL6_Unfamiliar.txt")

# Output summaries to the file
summary(model.overallsusc)
# Save
sink()

```

Familiar virus, HostStrain BALB/c vs C57BL/6: t = 2.774  p = 0.024941
Unfamiliar virus, HostStrain BALB/c vs C57BL/6: t = -1.495 p = 0.1763

BALB/c mice get significantly sicker than BL6 with the familiar virus, No major difference between mouse strains when infected with unfamliar virus. This is a bit meaningless though.



Aaaanyway, let's look at the effect of pooling treatment.

###### 4) Treatment effect. 1X vs 4X

```{r POOLING 1X vs 4X TESTS}
#Remove data that is not relevant
treattest <- subset(alldata, Treatment !="STOCK" & Treatment != "2X" & Sex == "F" & Treatment != "1XP")

# Subset to specify comparisons

 #treattest <- subset(treattest, Host == "BL6F")
 #treattest <- subset(treattest, HostStrain == "C57BL/6")
 treattest <- subset(treattest, Familiarity == "UNF")

#Level on 1X

treattest$Treatment <- relevel(factor(treattest$Treatment), ref = "1X")

#Scale and center.
treattest$Day <- scale(treattest$Day, center = TRUE, scale = TRUE)

#Run model
model.treat <- lmer(Pct ~ Day * Treatment + (Day|Line/ID), data = treattest, REML = TRUE)

# Open the file for writing
sink("LMM-Tradeoff-1X-vs-4X.txt")

# Output summaries to the file
summary(model.treat)
# Save
sink()

```

    1X vs 4X for:
Overall: t = -0.813 p = 0.443 
BALB adapted virus in fam and unfam: t = -0.906 p= 0.448   
BL6 adapted virus in fam and unfam: t =-0.445 p = 0.697
BALB/c mice infected with both viruses: t = -0.363 p = 0.727
C57BL/6 mice infected with both viruses: t = -0.996 p = 0.357
Familiar: t = -0.904 p = 0.396    
Unfamiliar: t = -0.522 p = 0.61925

No significant difference anywhere.

###### 5) Treatment effect. 1X vs 1XP

```{r POOLING 1X vs 1XP tests}
#Remove data that is not relevant
poolingtest <- subset(alldata, Treatment !="STOCK" & Treatment != "2X" & Sex == "F" & Treatment != "4X")

# Subset to specify comparisons

 #poolingtest <- subset(poolingtest, Host == "BL6F")
 #poolingtest <- subset(poolingtest, HostStrain == "BALB/C")
 poolingtest <- subset(poolingtest, Familiarity == "UNF")

#Level on 1X

poolingtest$Treatment <- relevel(factor(poolingtest$Treatment), ref = "1X")

#Scale and center.
poolingtest$Day <- scale(poolingtest$Day, center = TRUE, scale = TRUE)

#Run model
model.pooling <- lmer(Pct ~ Day * Treatment + Host + (Day|Line/ID), data = poolingtest, REML = TRUE)

# Open the file for writing
sink("LMM-Tradeoff-1X-vs-1XP_plus_host.txt")

# Output summaries to the file
summary(model.pooling)
# Save
sink()

```


```{r 1XP vs combined 1X 2X 4X}
#Remove data that is not relevant
poolingtest <- subset(alldata, Treatment !="STOCK" & Sex == "F")

# Subset to specify comparisons

 #poolingtest <- subset(poolingtest, Host == "BL6F")
 #poolingtest <- subset(poolingtest, HostStrain == "BALB/C")
 poolingtest <- subset(poolingtest, Familiarity == "UNF")

#Level on 1X

poolingtest$Treatment <- ifelse(poolingtest$Treatment %in% c("1X", "2X", "4X"), "1X2X4X", "1XP")

poolingtest$Treatment <- relevel(factor(poolingtest$Treatment), ref = "1X2X4X")

#Scale and center.
poolingtest$Day <- scale(poolingtest$Day, center = TRUE, scale = TRUE)

#Run model
model.pooling <- lmer(Pct ~ Day * Treatment + Host + (Day|Line/ID), data = poolingtest, REML = TRUE)

# Open the file for writing
sink("LMM-Tradeoff-1X2X4X-vs-1XP_plus_host.txt")

# Output summaries to the file
summary(model.pooling)
# Save
sink()

```

What about differences for overall virulence of all treatments, fam and unfam?

```{r OVERALL VIRULENCE DIFFERENCES}
#Remove data that is not relevant
poolingtest <- subset(alldata, Treatment !="STOCK" & Sex =="F")

# Subset to specify comparisons

 #poolingtest <- subset(poolingtest, Host == "BL6F")
 #poolingtest <- subset(poolingtest, HostStrain == "BALB/C")
 #poolingtest <- subset(poolingtest, Familiarity == "UNF")

#Level on 1X

poolingtest$Treatment <- relevel(factor(poolingtest$Treatment), ref = "1X")

#Scale and center.
poolingtest$Day <- scale(poolingtest$Day, center = TRUE, scale = TRUE)

#Run model
model.pooling <- lmer(Pct ~ Day * Treatment + Host + (Day|Line/ID), data = poolingtest, REML = TRUE)

# Open the file for writing
sink("LMM-ALL_Treatments_in_allHosts_Females_only.txt")

# Output summaries to the file
summary(model.pooling)
# Save
sink()

```

