---
title: "Model Test"
author: "RC"
date: "2024-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(dplyr)
```

Load data
```{r}
alldata <- read.csv("groupedvirulence7days.csv", sep=',', header=T)
```

First we look at the effect of passaged virus in each host compared to the ancestral virus in that host. No treatment comparison possible due to ancestral not having comparable treatments, so just an overall comparison.


Passaged vs unpassaged
```{r}
# Subset data
famtest <- subset(alldata, Familiarity == "FAM" & Treatment != "1XP" )

# Scale and center Day
famtest$Day <- scale(famtest$Day, center = TRUE, scale = TRUE)

# Hosts to iterate over
hosts <- c("BALBF", "BALBM", "BL6F")

# Open a connection to the output file
output_file <- "LMM_Results_Host_passaged_vs_unpassaged.txt"
sink(output_file)

cat("LMM results for ancestral vs evolved virus in each host")
# Iterate through each host as the reference level
for (host in hosts) {
  # Relevel Host
  famtest$Host <- relevel(factor(famtest$Host), ref = host)
  
  # Fit the model
  model <- lmer(Pct ~ Day * Status * Host + (Day | Line/ID), data = famtest, REML = TRUE)
  
  # Print header for clarity
  cat("\n\nResults with Host as Reference Level:", host, "\n")
  cat("---------------------------------------------------\n")
  
  # Output summary
  print(summary(model))
}

# Close the connection
sink()

```



For general virulence for each adapted strain, we will go with effect of Host only as there is no effect of treatment, so we keep the model simple.
```{r}
# Subset data
famtest <- subset(alldata, Familiarity == "FAM" & Treatment != "1XP" & Status == "Passaged")

# Scale and center Day
famtest$Day <- scale(famtest$Day, center = TRUE, scale = TRUE)

# Hosts to iterate over
hosts <- c("BALBF", "BALBM", "BL6F")

# Open a connection to the output file
output_file <- "LMM_Results_Host_Differences.txt"
sink(output_file)

cat("LMM results for weight loss comparisons between hosts [ Pct ~ Day * Host  + (Day | Line/ID)]")

# Iterate through each host as the reference level
for (host in hosts) {
  # Relevel Host
  famtest$Host <- relevel(factor(famtest$Host), ref = host)
  
  # Fit the model
  model <- lmer(Pct ~ Day * Host + (Day | Line/ID), data = famtest, REML = TRUE)
  
  # Print header for clarity
  cat("\n\nResults with Host as Reference Level:", host, "\n")
  cat("---------------------------------------------------\n")
  
  # Output summary
  print(summary(model))
}

# Close the connection
sink()

```


