---
title: "Model Test"
author: "RC"
date: "2024-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(dplyr)
```

Load data
```{r}
alldata <- read.csv("groupedvirulence7days.csv", sep=',', header=T)
```

First we look at the effect of passaged virus in each host compared to the ancestral virus in that host. No treatment comparison possible due to ancestral not having comparable treatments, so just an overall comparison.


Passaged vs unpassaged
```{r}
# Subset data
famtest <- subset(alldata, Familiarity == "FAM" & Treatment != "1XP" )

# Scale and center Day
famtest$Day <- scale(famtest$Day, center = TRUE, scale = TRUE)

# Hosts to iterate over
hosts <- c("BALBF", "BALBM", "BL6F")

# Open a connection to the output file
output_file <- "LMM_Results_Host_passaged_vs_unpassaged.txt"
sink(output_file)

cat("LMM results for ancestral vs evolved virus in each host")
# Iterate through each host as the reference level
for (host in hosts) {
  # Relevel Host
  famtest$Host <- relevel(factor(famtest$Host), ref = host)
  
  # Fit the model
  model <- lmer(Pct ~ Day * Status * Host + (Day | Line/ID), data = famtest, REML = TRUE)
  
  # Print header for clarity
  cat("\n\nResults with Host as Reference Level:", host, "\n")
  cat("---------------------------------------------------\n")
  
  # Output summary
  print(summary(model))
}

# Close the connection
sink()

```

Uncorrelated day and line slopes
```{r}

# Subset data
famtest <- subset(alldata, Familiarity == "FAM" & Treatment != "1XP" )

# Create a scaled Day column without overwriting the original
famtest$Day_z <- as.numeric(scale(famtest$Day, center = TRUE, scale = TRUE))

hosts <- c("BALBF", "BALBM", "BL6F")

output_file <- "LMM_Results_Host_passaged_vs_unpassaged_refit.txt"
sink(output_file)

cat("LMM results for passaged vs unpassaged, with Host releveling\n")
cat("Model: Pct ~ Day_z * Status * Host + random effects\n")
cat("Unit IDs: ID nested within Line\n\n")

for (host in hosts) {

  # Work on a local copy so you do not mutate famtest across iterations
  df <- famtest
  df$Host <- relevel(factor(df$Host), ref = host)

  cat("\n\nResults with Host as reference level: ", host, "\n", sep = "")
  cat("---------------------------------------------------\n")

  # Model 1: correlated random intercept/slope (may be singular)
  m_corr <- lmer(Pct ~ Day_z * Status * Host + (Day_z | Line/ID),
                 data = df, REML = TRUE)

  cat("\nModel with correlated random effects: (Day_z | Line/ID)\n")
  cat("isSingular: ", isSingular(m_corr), "\n", sep = "")
  cat("AIC: ", AIC(m_corr), "\n", sep = "")
  print(summary(m_corr))

  # Model 2: uncorrelated random intercept/slope (often resolves singularity)
  m_uncorr <- lmer(Pct ~ Day_z * Status * Host + (Day_z || Line/ID),
                   data = df, REML = TRUE)

  cat("\nModel with uncorrelated random effects: (Day_z || Line/ID)\n")
  cat("isSingular: ", isSingular(m_uncorr), "\n", sep = "")
  cat("AIC: ", AIC(m_uncorr), "\n", sep = "")
  print(summary(m_uncorr))
}

sink()
```


For general virulence for each adapted strain, we will go with effect of Host only as there is no effect of treatment, so we keep the model simple.
```{r}
# Subset data
famtest <- subset(alldata, Familiarity == "FAM" & Treatment != "1XP" & Status == "Passaged")

# Scale and center Day
famtest$Day <- scale(famtest$Day, center = TRUE, scale = TRUE)

# Hosts to iterate over
hosts <- c("BALBF", "BALBM", "BL6F")

# Open a connection to the output file
output_file <- "LMM_Results_Host_Differences.txt"
sink(output_file)

cat("LMM results for weight loss comparisons between hosts [ Pct ~ Day * Host  + (Day | Line/ID)]")

# Iterate through each host as the reference level
for (host in hosts) {
  # Relevel Host
  famtest$Host <- relevel(factor(famtest$Host), ref = host)
  
  # Fit the model
  model <- lmer(Pct ~ Day * Host + (Day | Line/ID), data = famtest, REML = TRUE)
  
  # Print header for clarity
  cat("\n\nResults with Host as Reference Level:", host, "\n")
  cat("---------------------------------------------------\n")
  
  # Output summary
  print(summary(model))
}

# Close the connection
sink()

```

Uncorrelated!

```{r}
# Create scaled Day as a numeric column (do not overwrite original Day)
famtest$Day_z <- as.numeric(scale(famtest$Day, center = TRUE, scale = TRUE))

hosts <- c("BALBF", "BALBM", "BL6F")

output_file <- "LMM_Results_Host_Differences_uncorr.txt"
sink(output_file)

cat("LMM results for weight loss comparisons between hosts\n")
cat("Model: Pct ~ Day_z * Host + (Day_z || Line/ID)\n")
cat("Random effects: uncorrelated random intercepts and slopes for Day_z\n")
cat("DF: Satterthwaite (lmerTest)\n\n")

for (host in hosts) {

  # Use a local copy so you do not mutate famtest across iterations
  df <- famtest
  df$Host <- relevel(factor(df$Host), ref = host)

  model <- lmer(Pct ~ Day_z * Host + (Day_z || Line/ID),
                data = df, REML = TRUE)

  cat("\n\nResults with Host as reference level: ", host, "\n", sep = "")
  cat("---------------------------------------------------\n")
  cat("isSingular: ", isSingular(model), "\n", sep = "")
  print(summary(model))
}

sink()
```


