---
title: "DISTANCE MATRIX MAKER"
author: "RC"
date: "2023-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
```

This script allows for obtaining distance matrices from pdb files, and fixing the names for the format A.HA.001.MET.

To get not just the default PDB but all chains go to on the PDB page -> Download Files -> Biological Assembly 1 (PDB - gz)

Notes: To make this run for all pdbs not much has to be changed:
  Check names/paths
  Check valid chains to keep
  Check starting positions for each chain (see which aminoacid each chain starts in to properly assign aminoacid position    and names)

The last option "No threshold" is the best for centroids.

```{r}
python_path <- normalizePath("C:/Users/rsk/AppData/Local/Programs/Python/Python312/python.exe")
reticulate::use_python(python_path)
```


```{r}
# Install necessary Python packages
reticulate::py_install(c("biopython", "numpy", "pandas"))

```


```{python}
# Install necessary Python packages
!pip install biopython numpy pandas

```

###MAKE DISTANCE MATRIX####

Try to get distance matrix for pdb
```{python}
# Import necessary libraries
from Bio.PDB import PDBParser
import numpy as np
import pandas as pd

# Define the calculate_distance_matrix function
def calculate_distance_matrix(structure):
    atom_list = list(structure.get_atoms())
    num_atoms = len(atom_list)
    distance_matrix = np.zeros((num_atoms, num_atoms))

    for i in range(num_atoms):
        for j in range(i + 1, num_atoms):
            distance_matrix[i, j] = distance_matrix[j, i] = atom_list[i] - atom_list[j]

    return distance_matrix

# Define the main function
def main(pdb_filename, output_csv):
    parser = PDBParser(QUIET=True)
    structure = parser.get_structure('protein', pdb_filename)

    distance_matrix = calculate_distance_matrix(structure)

    # Convert the distance matrix to a Pandas DataFrame
    df_distance_matrix = pd.DataFrame(distance_matrix)

    # Save the DataFrame to a CSV file
    df_distance_matrix.to_csv(output_csv, index=False)

# Specify the PDB filename and the output CSV filename
pdb_filename = "4oph.pdb"
output_csv = "distance_matrix_NS1.csv"

main(pdb_filename, output_csv)

```

Try below 50 angstrom, with info on residue:
```{python}
# Import necessary libraries
from Bio.PDB import PDBParser
import numpy as np
import pandas as pd

# Define the calculate_distance_matrix function
def calculate_distance_matrix(structure):
    atom_list = list(structure.get_atoms())
    num_atoms = len(atom_list)
    distance_matrix = np.zeros((num_atoms, num_atoms))

    for i in range(num_atoms):
        for j in range(i + 1, num_atoms):
            distance = atom_list[i] - atom_list[j]

            # Filter distances below 50 Angstroms
            if distance < 50:
                distance_matrix[i, j] = distance_matrix[j, i] = distance

    return distance_matrix

# Define the main function
def main(pdb_filename, output_csv):
    parser = PDBParser(QUIET=True)
    structure = parser.get_structure('protein', pdb_filename)

    distance_matrix = calculate_distance_matrix(structure)

    # Extract information for row and column names
    names = [
        f"{atom.get_full_id()[2]}_{structure.id}_{atom.get_full_id()[3][1]:03d}_{atom.get_parent().get_resname()}"
        for atom in structure.get_atoms()
    ]

    # Convert the distance matrix to a Pandas DataFrame with modified names
    df_distance_matrix = pd.DataFrame(distance_matrix, index=names, columns=names)

    # Save the DataFrame to a CSV file
    df_distance_matrix.to_csv(output_csv)

# Specify the PDB filename and the output CSV filename
pdb_filename = "6nhq.pdb"
output_csv = "distance_matrix_6nhq.csv"

# Call the main function
main(pdb_filename, output_csv)

```

And with centroids (may mess up names)
```{python}
# Import necessary libraries
from Bio.PDB import PDBParser
import numpy as np
import pandas as pd

# Define the calculate_centroid_distances function
def calculate_centroid_distances(structure):
    # Dictionary to store centroid coordinates for each residue
    centroids = {}

    # Iterate through all residues
    for model in structure:
        for chain in model:
            for residue in chain:
                # Calculate centroid coordinates as the average of all atom coordinates in the residue
                centroid = np.mean([atom.get_coord() for atom in residue], axis=0)
                # Extract relevant information for the key
                chain_id = chain.id
                residue_number = residue.id[1]
                residue_name = residue.get_resname()
                key = (chain_id, model.id, residue_number, residue_name)
                centroids[key] = centroid

    # Calculate distances between centroids
    num_residues = len(centroids)
    centroid_distances = np.zeros((num_residues, num_residues))

    keys = list(centroids.keys())

    for i in range(num_residues):
        for j in range(i + 1, num_residues):
            distance = np.linalg.norm(centroids[keys[i]] - centroids[keys[j]])
            centroid_distances[i, j] = centroid_distances[j, i] = distance

    return centroids, centroid_distances

# Define the main function
def main(pdb_filename, output_csv):
    parser = PDBParser(QUIET=True)
    structure = parser.get_structure('protein', pdb_filename)

    centroids, centroid_distances = calculate_centroid_distances(structure)

    # Convert the centroid distance matrix to a Pandas DataFrame with modified names
    names = [f"{key[0]}_{structure.id}_{key[2]:03d}_{key[3]}" for key in centroids.keys()]
    df_centroid_distances = pd.DataFrame(centroid_distances, index=names, columns=names)

    # Save the DataFrame to a CSV file
    df_centroid_distances.to_csv(output_csv)


# Specify the PDB filename and the output CSV filename
pdb_filename = "6nhq.pdb"
output_csv = "distance_matrix_6nhq.csv"

# Call the main function
main(pdb_filename, output_csv)

```

Below 50A, Correct to not put zeroes when it's above 50 A (it's empty instead)

```{python}
# Import necessary libraries
from Bio.PDB import PDBParser
import numpy as np
import pandas as pd

# Define the calculate_centroid_distances function
def calculate_centroid_distances(structure, distance_threshold=50):
    # Dictionary to store centroid coordinates for each residue
    centroids = {}

    # Iterate through all residues
    for model in structure:
        for chain in model:
            for residue in chain:
                # Calculate centroid coordinates as the average of all atom coordinates in the residue
                centroid = np.mean([atom.get_coord() for atom in residue], axis=0)
                # Extract relevant information for the key
                chain_id = chain.id
                residue_number = residue.id[1]
                residue_name = residue.get_resname()
                key = (chain_id, model.id, residue_number, residue_name)
                centroids[key] = centroid

    # Calculate distances between centroids
    num_residues = len(centroids)
    centroid_distances = np.zeros((num_residues, num_residues))  # Initialize with zeros

    keys = list(centroids.keys())

    for i in range(num_residues):
        for j in range(i + 1, num_residues):
            distance = np.linalg.norm(centroids[keys[i]] - centroids[keys[j]])

            # Check if the distance is above the threshold
            if distance > distance_threshold:
                centroid_distances[i, j] = centroid_distances[j, i] = np.nan
            else:
                centroid_distances[i, j] = centroid_distances[j, i] = distance

    return centroids, centroid_distances


# Define the main function
def main(pdb_filename, output_csv):
    parser = PDBParser(QUIET=True)
    structure = parser.get_structure('protein', pdb_filename)

    centroids, centroid_distances = calculate_centroid_distances(structure)

    # Convert the centroid distance matrix to a Pandas DataFrame with modified names
    names = [f"{key[0]}_{structure.id}_{key[2]:03d}_{key[3]}" for key in centroids.keys()]
    df_centroid_distances = pd.DataFrame(centroid_distances, index=names, columns=names)

    # Save the DataFrame to a CSV file
    df_centroid_distances.to_csv(output_csv)

# Specify the PDB filename and the output CSV filename
pdb_filename = "4oph1.pdb"
output_csv = "distance_matrix_NS1b.csv"

# Call the main function
main(pdb_filename, output_csv)

```

No threshold (Centroids)
```{python}
# Import necessary libraries
from Bio.PDB import PDBParser
import numpy as np
import pandas as pd

# Define the calculate_centroid_distances function
def calculate_centroid_distances(structure):
    # Dictionary to store centroid coordinates for each residue
    centroids = {}

    # Iterate through all residues
    for model in structure:
        for chain in model:
            for residue in chain:
                # Calculate centroid coordinates as the average of all atom coordinates in the residue
                centroid = np.mean([atom.get_coord() for atom in residue], axis=0)
                # Extract relevant information for the key
                chain_id = chain.id
                residue_number = residue.id[1]
                residue_name = residue.get_resname()
                key = (chain_id, model.id, residue_number, residue_name)
                centroids[key] = centroid

    # Calculate distances between centroids
    num_residues = len(centroids)
    centroid_distances = np.zeros((num_residues, num_residues))

    keys = list(centroids.keys())

    for i in range(num_residues):
        for j in range(i + 1, num_residues):
            distance = np.linalg.norm(centroids[keys[i]] - centroids[keys[j]])
            centroid_distances[i, j] = centroid_distances[j, i] = distance

    return centroids, centroid_distances

# Define the main function
def main(pdb_filename, output_csv):
    parser = PDBParser(QUIET=True)
    structure = parser.get_structure('protein', pdb_filename)

    centroids, centroid_distances = calculate_centroid_distances(structure)

    # Convert the centroid distance matrix to a Pandas DataFrame with modified names
    names = [f"{key[0]}_{structure.id}_{key[2]:03d}_{key[3]}" for key in centroids.keys()]
    df_centroid_distances = pd.DataFrame(centroid_distances, index=names, columns=names)

    # Save the DataFrame to a CSV file
    df_centroid_distances.to_csv(output_csv)

# Specify the PDB filename and the output CSV filename
pdb_filename = "6qnw.pdb"
output_csv = "distance_matrix_6QNW.csv"

# Call the main function
main(pdb_filename, output_csv)

```


####FIX NAMES###

This fixes naming format if I need to shift numbering. 

BUT BEFORE MUST FIX CSV TO MAKE SURE THE CHAINS HAVE THE RIGHT NAMES (go to excel and just rename series of aminoacids appropriately)

Remember to add "Residue" to the empty cell at the end.

___NOTICE: I modified the HA distance matrix to conform to the uncleaved numbering (G234W instead of G218W). Alternatively the original numbering can be maintained here but then it has to be changed in the column names of the ancestral/derived counts file. Either way works___


```{r}
# Assuming your data frame is named 'df'
df <- read.csv("distance_matrix_6QNW.csv", row.names = 1)

# SELECT CHAINS TO KEEP: Define a vector of valid chain identifiers (A-F)

valid_chains <- c("A", "B", "C", "D", "E", "F")

# 1: Delete rows and columns that don't start with A-F
df <- df[grepl("^(A|B|C|D|E|F)_", rownames(df)) & grepl("^(A|B|C|D|E|F)_", colnames(df)), ]

# 2: Define starting positions for each chain
starting_positions <- c(A = 1, B = 1)

# 3: Update row and column names
for (chain in valid_chains) {
  chain_rows <- grepl(paste0("^", chain, "_"), rownames(df))
  chain_cols <- grepl(paste0("^", chain, "_"), colnames(df))
  
  if (any(chain_rows) && any(chain_cols)) {
    starting_position <- starting_positions[chain]
    
    # Generate new names with updated numbering and the original amino acid names
    new_row_numbers <- starting_position + seq_len(sum(chain_rows)) - 1
    new_row_names <- paste0(chain, "_NS1_", sprintf("%03d", new_row_numbers), "_", substring(rownames(df)[chain_rows], nchar(rownames(df)[chain_rows])-2))
    
    new_col_numbers <- starting_position + seq_len(sum(chain_cols)) - 1
    new_col_names <- paste0(chain, "_NS1_", sprintf("%03d", new_col_numbers), "_", substring(colnames(df)[chain_cols], nchar(colnames(df)[chain_cols])-2))
    
    rownames(df)[chain_rows] <- new_row_names
    colnames(df)[chain_cols] <- new_col_names
  }
}

#Change separators into periods.
# Replace all underscores with dots in row and column names
rownames(df) <- gsub("_", ".", rownames(df))
colnames(df) <- gsub("_", ".", colnames(df))


# 4: Delete rows and columns containing "HOH" or "NAG"
rows_to_delete <- grepl("HOH|NAG", rownames(df))
cols_to_delete <- grepl("HOH|NAG", colnames(df))

# Exclude rows and columns with names containing "HOH" or "NAG"
df2 <- df[!rows_to_delete, !cols_to_delete]

# 5: Save the updated dataframe to a new CSV file
write.csv(df2, "6QNW-dm-nolimit.csv")

```


Without renumbering
```{r}
# Assuming your data frame is named 'df'
df <- read.csv("distance_matrix_6QNW.csv", row.names = 1)

# SELECT CHAINS TO KEEP: Define a vector of valid chain identifiers (A-F)

valid_chains <- c("A", "B", "C", "D", "E", "F")

# 1: Delete rows and columns that don't start with A-F
df <- df[grepl("^(A|B|C|D|E|F)_", rownames(df)) & grepl("^(A|B|C|D|E|F)_", colnames(df)), ]

#Change separators into periods.
# Replace all underscores with dots in row and column names
rownames(df) <- gsub("_", ".", rownames(df))
colnames(df) <- gsub("_", ".", colnames(df))


# 4: Delete rows and columns containing "HOH" or "NAG"
rows_to_delete <- grepl("HOH|NAG", rownames(df))
cols_to_delete <- grepl("HOH|NAG", colnames(df))

# Exclude rows and columns with names containing "HOH" or "NAG"
df2 <- df[!rows_to_delete, !cols_to_delete]

# 5: Save the updated dataframe to a new CSV file
write.csv(df2, "6QNW-dm-nolimit.csv")

```

